### 2019-12-10 ms/ms.Rmd
version submitted to Frontiers

---
title: "A stomatal model of anatomical tradeoffs between photosynthesis and pathogen defense"
runtitle: "Stomata tradeoff photosynthesis for defense"
documentclass: frontiersSCNS
author:
  - name: Christopher D. Muir
    affiliation: '1'
    etal: Muir
    email: cdmuir@hawaii.edu
affiliation:
  - id: '1'
    department: School of Life Sciences
    institution: University of Hawaii
    city: Honolulu
    state: Hawaii
    country: USA
output: 
  rticles::frontiers_article:
    includes:
      in_header: header.tex
csl: frontiers.csl
bibliography: stomata-tradeoff.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\begin{abstract}

Stomatal pores control both leaf gas exchange and are an entry for many plant pathogens, setting up the potential for tradeoffs between photosynthesis and defense. To prevent colonization and limit infection, plants close their stomata after recognizing pathogens. In addition to closing stomata, anatmoical shifts to lower stomatal density and/or size may also limit pathogen colonization, but such developmental changes would permanently reduce the gas exchange capacity for the life of the leaf. I developed and analyzed a spatially explicit model of pathogen colonization on the leaf as a function of stomatal size and density, anatomical traits which determine maximum rates of gas exchange. The model predicts greater stomatal size or density increases colonization, but the effect is most pronounced when stomatal cover is low. I also derived scaling relationships between stomatal size and density that preserves a given probability of colonization. These scaling relationships set up a potential conflict between maximizing defense and minimizing stomatal cover. To my knowledge, this is the first mathematical model connecting gas exchange and pathogen defense via stomatal anatomy. It makes predictions that can be tested with experiments and may explain variation in stomatal anatomy among plants. The model is generalizable to many types of pathogens, but lacks significant biological realism that may be needed for accurate predictions.

\tiny
 \keyFont{ \section{Keywords:} anatomy, leaf gas exchange, model, pathogen, photosynthesis, scaling, stomata, tradeoff }

\end{abstract}

# Introduction {-}

<!-- overview -->
Stomata evolved to regulate gas exchange in and out of the leaf [@hetherington_role_2003; @berry_stomata:_2010; @chater_origins_2017], but many plant pathogens take advantage of these chinks in the leaf cuticular armor to infect prospective hosts [@zeng_plant_2010; @mclachlan_gate_2014; @melotto_stomatal_2017]. The density and size of stomata set the anatomical maximum rate of stomatal conductance to CO$_2$ and water vapor [@brown_static_1900; @parlange_stomatal_1970; @franks_effect_2001; @franks_maximum_2009; @lehmann_effects_2015; @sack_developmental_2016; @harrison_influence_2019], but the pore area shrinks and expands in response to internal and external factors to regulate gas exchange dynamically [@buckley_how_2019]. Many plant pathogens, including viruses [@murray_plant_2016], bacteria [@melotto_plant_2006; @underwood_role_2007], protists [@fawke_oomycete_2015], and fungi [@hoch_signaling_1987; @zeng_plant_2010] use stomatal pores to gain entry into the leaf. Since stomatal conductance is a major limitation on photosynthesis [@farquhar_stomatal_1982; @jones_partitioning_1985] while pathogens reduce fitness, this sets up a potential tradeoff between increased photosynthesis and defense against pathogens. Although there have been many empirical studies on the effect that pathogens have on stomata, there is no theoretical framework in which to place these findings. Lack of a theoreatical framework makes it difficult to answer general questions about how selection for pathogen defense constrains maximum rates of gas exchange.

<!-- paragraph on stomatal anatomy and gsmax -->
Stomatal anatomy is the key link between gas exchange and pathogen colonization. The density and size of stomata not only determines the theoretical maximum stomatal conductance (\gsmax), but is also proportional to the operational stomatal conductance (\gsop) in many circumstances [@franks_plasticity_2009; @dow_integrated_2014; @franks_new_2014; @mcelwain_using_2016; @murray_consistent_2019]. Therefore, I use anatomical \gsmax{} as a proxy for \gsop{} and do not address dynamic changes in stomatal aperture (see [Discussion]). @harrison_influence_2019 recently reviewed the relationship between stomatal anatomy and gas exchange in detail.

<!-- paragraph on stomatal anatomy and pathogen defense -->
Many pathogens rely on stomata to gain entry into the leaf, but it is unclear how anatomical traits that determine \gsmax{} (size and density) affect the ability of pathogens to colonize leaves. The impact of pathogens on host fitness is complex, but after a pathogen reaches a host, the first major step is colonization. Once infected, a pathogen can reduce fitness. Susceptible hosts can lose much of their biomass or die, but even resistant hosts must allocate resources to defense or reduce photosynthesis because of necrosis around sites of infection. Plants can limit colonization physiologically by closing stomata after they recognize pathogens, called stomatal defense [@melotto_stomatal_2017]. Anatomy may be another layer of defense; plants can reduce pathogen colonization by developing leaves with lower stomatal density and/or size. Infection increases in leaves with higher stomatal density [@mckown_association_2014; @tateda_host_2019; @dutton_bacterial_2019; @fetter_trade-offs_2019], which suggests that both anatomical and physiological responses (stomatal closure) affect host colonization. By the same logic, if stomatal density were held constant, larger stomata should also increase colonization because they occupy more area. One key difference between physiological and anatomical defenses is that stomatal closure reduces gas exchange transiently, whereas anatomy constrains gas exchange throught the life of the leaf. Because we do not understand the relationship between stomatal anatomy and colonization well, we cannot predict the relationship between \gsmax{} and infection.

<!-- paragraph on size-density scaling? -->
If stomatal size and density affect pathogen colonization, selection to limit colonization may shape stomatal size-density scaling relationships. Botanists have long recognized that stomatal size and density are inversely correlated [@weiss_untersuchungen_1865; @ticha_photosynthetic_1982; @hetherington_role_2003; @sack_hydrology_2003; @franks_co_2009; @brodribb_unified_2013; @de_boer_optimal_2016], but the evolutionary origin of this relationship is not yet known. Here I argue that pathogens could shape selection on this relationship. Explanations for inverse size-density scaling are usually cast in terms of preserving \gsmax{} and/or total epidermal area allocated to stomata [\fs; @de_boer_optimal_2016] because there are many combinations of stomatal size and density that have same \gsmax{} or same \fs{}:

\begin{align}
  g_\text{s,max} & = b m D S ^ {0.5} \label{eq:gsmax} \\
  f_\text{S} & = D S \label{eq:fs}.
\end{align}

$b$ and $m$ are assumed to be biophysical and morphological constants, *sensu* @sack_developmental_2016 (see [Supplementary Material]). If size and density also affect pathogen colonization, then selection from foliar pathogens could significantly alter the size-density scaling relationship. The empirical size-density scaling relationship is linear on a log-log scale, determined by an intercept $\alpha$ and slope $\beta$:

\begin{align}
  D & = e ^ \alpha S ^ {-\beta}; \\
  d & = \alpha - \beta S.
\end{align}

For brevity, $d = \text{log}(D)$ and $s = \text{log}(S)$. Rearranging Equations \ref{eq:gsmax} and \ref{eq:fs}, a scaling relationship where $\beta = 0.5$ preserves \gsmax{} while $\beta = 1$ preserves \fs.

How would adding pathogens alter these predicted scaling relationships? For simplicity, imagine two environments, one without foliar pathogens and one with lots. In the absence of foliar pathogens, we expect size-density scaling to preserve \gsmax{}, \fs{}, or some least-cost combination of them. What happens when we introduce pathogens? Assuming that stomatal size and density increase pathogen colonization, then selection will favor reduced size and/or density. This would change the intercept $\alpha$ but not the slope. The effect of foliar pathogens on the slope depends on the relationship between size, density, and probability of colonization. If the probability of colonization is proportional to the product of *linear* stomatal size ($S ^ {0.5}$) and density ($\propto D S ^ {0.5}$ as for \gsmax) then it has the same effect on the slope as \gsmax{} because there are many combinations of $D$ and $S ^ {0.5}$ that have same probability of colonization. If the probability of colonization is proportional to the product of *areal* stomatal size ($S$) and density ($\propto D S$ as for \fs) then it has the same effect on the slope as \fs{} because there are many combinations of $D$ and $S$ that have same probability of colonization. Alternatively, the probability of colonization may have a different scaling relationsip (neither 0.5 nor 1) or may be nonlinear on a log-log scale. Unlike \gsmax{} and \fs{}, we do not have theory to predict a stomatal size-density relationship that preserves the probability of colonization.

<!-- paragraph on what we don't know -->
In summary, the physical relationship between stomatal size, density, and conductance is well established [@harrison_influence_2019]. The same traits likely affect the probability of pathogen colonization, but we do not have a theoretical model that makes quantitative predictions. The inverse stomatal size-density relationship has usually been explained in terms of preserving stomatal conductance and/or stomatal cover, but selection by pathogens might alter scaling. To address these gaps, the goals of this study are to 1) introduce a spatially explicit model pathogen colonization on the leaf surface; 2) use the model to predict the relationship between \gsmax, \fs{}, and the probability of colonization; 3) work out what these relationships predict about stomatal size-density scaling.

# Model {-}

In this section, I introduce a spatially explicit model of pathogen colonization on a leaf surface. I explain the model structure and assumptions here; the [Materials and Methods] section below describes how I analyzed the model to address the goals of the study. For generality, I refer to a generic ``pathogen'' that lands on leaf and moves to a stomate. The model is agnostic to the type of pathogen (virus, bacterium, fungus, etc.) and the specific biological details of how it moves (biotrophy). For example, motile bacterial cells can land and move around where fungi may germinate from a cyst and grow until they form an appresorium for infection. These very different biotropic movements on the leaf are treated identically here. I used Sympy version 1.3 [@meurer_sympy:_2017] for symbolic derivations.

**Table 1** | Glossary of mathematical symbols. The columns indicate the mathematical Symbol used in the paper, the associated symbol used in R scripts,  scientific Units, and a verbal Description.

 Symbol | R      | Units  | Description
--------|--------|--------|-----------------------------------------------------
 $D,d$              | `D, d`    | $\textrm{mm}^{-2}$    | stomatal density ($d = \text{log}~D$)
 $f_s$              | `f_s`     | none                  | stomatal cover ($f_s = D S$)
 $g_\text{s,max}$   | `g_smax`  | $\text{mol}~\text{m}^{-2}~\text{s}^{-1}$ | anatomical maximum stomatal conductance 
 $g_\text{s,op}$    | `g_sop`   | $\text{mol}~\text{m}^{-2}~\text{s}^{-1}$ | operational stomatal conductance 
 $H$                | `H`       | $\mu\textrm{m} ^{-1}$ | death rate of pathogen on leaf surface
 $R$                | `R`       | $\mu\textrm{m}$       | stomatal radius ($S = 2 \pi R ^ 2$)
 $S,s$              | `S, s`    | $\mu\textrm{m}^2$     | stomatal size ($s = \text{log}~S$)
 $\theta_i$         | `theta_i` | radians               | angles between pathogen ($x_p,y_p$) and lines tangent to the circumfrence of stomate $i$
 $U$                | `U`       | $\mu\textrm{m}$       | interstomatal distance
 $v_i$              | `v_i`     | $\mu\textrm{m}$       | distance between pathogen ($x_p,y_p$) and stomate $i$
 $x_i,y_i$          | `x_i,y_i` | $\mu\textrm{m}$       | position of stomate $i$
 $x_p,y_p$          | `x_p,y_p` | $\mu\textrm{m}$       | starting position of pathogen

## Spatial representation of stomata {-}

Stomata develop relatively equal spacing to minimize resistance to lateral diffusion [@morison_lateral_2005], allow space between stomata [@dow_physiological_2014], and prevent stomatal interference [@lehmann_effects_2015]. Here I assume that stomata are arrayed in an equilateral triangular grid with a density $D$ and size (area) $S$. This assumption ignores veins, trichomes, and within-leaf variation in stomatal density. Stomata are therefore arrayed in an evenly spaced grid (Figure \ref{fig:fig1}a). The interstomatal distance $U$, measured as the distance from the center of one stomata to the next, is the maximal diagonal of the hexagon in $\mu\textrm{m}$ that forms an equal area boundary between neighbording stomata. The area of a hexagon is $A_\textrm{hexagon} = \frac{\sqrt{3}}{2} U ^ 2$. By definition the stomatal density is the inverse of this area, such that $D = A^{-1}_\textrm{hexagon} = \frac{2}{\sqrt{3}} U ^ {-2}$. Therefore, interstomatal distance can be derived from the stomatal density as:

$$ D = \frac{2}{\sqrt{3}} U ^{-2} $$
$$ U = \bigg(\frac{2}{\sqrt{3}} D ^{-1}\bigg) ^ {0.5} $$

For example, if the density is $D = 10 ^ 2~\textrm{mm}^{-2} = 10 ^ {-4}~\mu\textrm{m}^{-2}$, then $U$ is `r round(sqrt(2 / (sqrt(3) * 1e-4)), 1)` $\mu \textrm{m}$. @parkhurst_diffusion_1994 described this result previously. I also make the simplifying assumption that stomata are perfectly circular with radius $R$. This may be approximately true for fully open stomata with kidney-shaped guard cells. Although I assume stomata are circular here, in calculating \gsmax, I assume typical allometric relationships between length, width, and pore area  (see [Supplementary Material]).

\begin{figure}
  \centering
	\includegraphics[width=6.5in]{../figures/fig1.jpeg}
	\caption{\textbf{A spatially explicit model of stomatal anatomy and pathogen colonization}. \textbf{a}. Stomata are assumed to be in a homogenous equilateral triangular grid, which means that we can extrapolate from \textbf{b.} a focal triangle to the entire leaf. The circles represent idealized stomata; the grey lines between them are for visualization. \textbf{c.} By symmetry, a single focal region within the focal triangle can be modeled and extrapolated to the rest of the area. \textbf{d.} The model assumes that a pathogen, depicted as a grey rod, lands somewhere on the leaf surface and will sucessfully locate a stomate if it moves at the correct angle, depicted by the grey polygons.}
	\label{fig:fig1}
\end{figure}

## Spatial representation of pathogen search {-}

Now imagine that a pathogen lands at a uniform random position within the focal region and must arrive at a stomate to colonize. If it lands on a stomate, then it infects the leaf with probability 1; if it lands between stomata, then it infects the leaf with probability $p_\text{locate}$. This is the probability that it locates a stomate, which I will derive below. The probabilities of landing on or between a stomate are \fs{} and $1 - f_S$, respectively. Hence, the total probability of colonization is:

\begin{equation}
  \label{eq:p_colonize}
  p_\text{colonize} = f_\text{S} + (1 - f_\text{S}) p_\text{locate}.
\end{equation}

I assume that the pathogen cannot sense where stomata are and orients at random, thereafter traveling in that direction. If it successfully locates a stomate, it colonizes the leaf, but otherwise does not infect. If there is a high density of stomata and/or large stomata, the probability of locating a stomate increases. By assuming that stomata form an equilateral triangular grid (see above), we can extrapolate what happens in a focal triangle (Figure \ref{fig:fig1}b) by symmetry. Further, since an equilateral triangle can be broken up into six identical units (Figure \ref{fig:fig1}c), we can simply calculate $p_\text{locate}$ in this focal area. This implicitly assumes that the probability of colonzing stomata outside the focal area is 0 because they are too far away.

Imagine that the pathogens lands in position ($x_p,y_p$) within the triangle. The centroid of the triangle is at position ($x_c,y_c$) and a reference stomate is at position $(0,0)$ (Figure \ref{fig:fig2}a). Therefore $x_c = U / 2$ and $y_c = \sqrt{3} U / 6$. The other stomata are at positions $(U/2, \sqrt{3} U /2)$ and $(U,0)$ (Figure \ref{fig:fig2}). $x_p$ and $y_p$ are defined as the horizontal and vertical distances, respectively, from the pathogen to the reference stomate at position $(0,0)$.

Given that the pathogen starts at position ($x_p,y_p$), what's the probability of contacting one of the stomata at the vertices of the focal triangle? I assume the probability of contacting a stomate is equal to the proportion of angular directions that lead to a stomate (Figure \ref{fig:fig1}d). I solved this by finding the angles ($\theta_1, \theta_2, \theta_3$) between lines that are tangent to the outside of the three stomata and pass through ($x_p,y_p$) (Figure \ref{fig:fig2}a). If stomate $i$ is centered at ($x_i,y_i$), the two slopes of tangency as function of pathogen position are:

\begin{align}
  t_{i,1}(x_p,y_p) & = \frac{-R e_{i,2}(x_p,y_p) + e_{i,3}(x_p,y_p)}{e_{i,1}(x_p,y_p)} \\
  t_{i,2}(x_p,y_p) & = \frac{R e_{i,2}(x_p,y_p) + e_{i,3}(x_p,y_p)}{e_{i,1}(x_p,y_p)}
\end{align}

where

\begin{align}
  e_{i,1}(x_p,y_p) & = (R^2 - x_i^2 + 2 x_i x_p - x_p^2), \\
  e_{i,2}(x_p,y_p) & = \sqrt{-e_{i,1} + (y_i - y_p) ^ 2}, \\
  e_{i,3}(x_p,y_p) & = - x_i y_i + x_i y_p + x_p y_i - x_p y_p.
\end{align}

Note that $i \in \{1, 2, 3\}$, indexing the three stomata in the focal triangle. The angle in radians between $t_{i,1}(x_p,y_p)$ and $t_{i,2}(x_p,y_p)$ is:

\begin{equation}
  \theta_i(x_p,y_p) = \text{arctan}\bigg(\frac{t_{i,1}(x_p,y_p) - t_{i,2}(x_p,y_p)}{1 + (t_{i,1}(x_p,y_p) t_{i,2}(x_p,y_p))}\bigg)
\end{equation}

\begin{figure}
\centering
\input{../figures/fig2.tex}
\caption{\textbf{Spatial representation of stomata and pathogen.} \textbf{a.} The pathogen starts at a uniform random position within the focal region denoted ($x_p,y_p$). Within the focal triangle, the reference stomata is at position (0,0) by definition, and other stomatal positions are determined by the interstomatal distance $U$. \textbf{b.} Within the focal region, a pathogen can land within the stomate (white cirlce with grey outline and radius $R$) or in the grey area. The outer borders of this area are shown and depend on $R$ and $U$. For a given position $x$, there is a minimum $y$-value ($y_\text{min}$, dashed line) and maximum $y$-value ($y_\text{max}$, solid line).}
\label{fig:fig2}
\end{figure}

I further assumed that the longer distance a pathogen must travel, the less likely it would be to locate a stomate. For example, if stomata are at very low density, then a pathogen may die before it reaches a stomate because of UV, dessication, or another factor. I included this effect by assuming the probability of reaching a stomate declines exponentially at rate $H$ with the Euclidean distance $v_i(x_p,y_p)$ between the pathogen location and the edge of stomata $i$, which is distance $R$ from its center at $x_i,y_i$:

\begin{equation}
  v_i(x_p,y_p) = \sqrt{(x_i - x_p) ^ 2 + (y_i - y_p) ^ 2} - R.
\end{equation}

The probability of locating a stomate as a function of $x_p$ and $y_p$ ($f_\text{locate}(x_p,y_p)$) is the sum of the angles divided by $2\pi$, discounted by their distance from the stomate:

\begin{equation}
  f_\text{locate}(x_p,y_p) = \frac{1}{2 \pi} \sum_{i=1}^3 e^{-H v_i(x_p,y_p)} \theta_i(x_p,y_p)
\end{equation}

When $H = 0$, $p_\text{locate}$ is the fraction of angles that lead from $(x_p,y_p)$ to a stomate. When $H > 0$, $p_\text{locate}$ is proportional to this fraction, but less than it depending on stomatal density, size, and starting location of the pathogen.

To obtain the average $p_\text{locate}$, we must integrate $f_\text{locate}(x_p,y_p)$ over all possible starting positions $(x_p,y_p)$ within the focal area. The focal area is a 30-60-90 triangle with vertices at the center of the reference stomate $(0,0)$, the midpoint of baseline $(U/2,0)$, and the centroid of the focal trinagle $(U/2, \sqrt{3} / 6 U)$ (Figure \ref{fig:fig1}c). Colonization occurs with probability 1 if the pathogen lands in the reference stomate, so we need to integrate the probability of colonization if it lands elsewhere. This region extends from the edge of the stomate, at $\sqrt{3} / 2 R$ to $U/2$ (Figure \ref{fig:fig2}b). At any $x$, we integrate from the bottom of the focal area ($y_\text{min}$) to the top ($y_\text{max}$):

\begin{align}
  y_\text{min} = f(x) & = \begin{cases}
    \sqrt{R ^ 2 - x ^ 2}, & \text{if } \frac{\sqrt{3}}{2} R < x < R \\ 
    0, & \text{if } R \leq x \leq \frac{U}{2} \end{cases} \\
  y_\text{max} = f(x) & = \frac{\sqrt{3}}{3} x
\end{align}

The integral is:

\begin{equation}
  p_\text{locate} = \frac{1}{a_\text{focal}}\int\limits_{\frac{\sqrt{3}}{2} R} ^ {U / 2} \int\limits_{y_\mathrm{min}} ^ {y_\mathrm{max}} f_\text{locate}(x, y) \,dx \,dy
\end{equation}

$a_\text{focal}$ is the area of the focal region depicted in grey in Figure \ref{fig:fig2}b:

$$ a_\text{focal} = \frac{U ^ 2}{8 \sqrt{3}} - \frac{\pi R ^ 2}{12} $$

# Materials and Methods {-}

The [Model] calculates a probability of host colonization (Equation \ref{eq:p_colonize}) as a function stomatal density, size, and position of a pathogen on the leaf. I solved $p_\text{colonize}$ using the `integral2()` function in the **pracma** package version 2.2.5 [@borchers_pracma:_2019] for numerical integration. I used R version 3.6.1 [@r_core_team_r:_2019] for all analyses and wrote the paper in **rmarkdown** version 1.17 [@xie_r_2018; @allaire_rmarkdown:_2019]. Source code is deposited on GitHub (https://github.com/cdmuir/stomata-tradeoff) and will be arcived on Zenodo upon publication.

## What is the relationship between stomatal size, density, and colonization? {-}

I calculated $p_\text{colonize}$ over a biologically plausible grid of stomatal size and density for hypostomatous species based on @de_boer_optimal_2016. Stomatal density ranges from $10^1 - 10^{3.5}~\text{mm}^{-2}$; stomatal size ranges from $10^1 - 10^{3.5}~\mu\text{m}^2$. I only considered combinations of size and density where \fs{} was less than 1/3. For simplicity, I have not extended the current analysis to amphistomatous leaves. I crossed stomatal traits with three levels of $H \in \{0, 0.01, 0.1\}$. When $H = 0$, a pathogen perists indefinitely on the leaf surface. $H = 0.01$ and $H = 0.1$ correspond to low and high death rates, respectively. These values are not necessarily realistic, but illustrate qualitatively how a hostile environment on the leaf surface alters model predictions.

## How do pathogens alter optimal stomatal size-density scaling? {-}

The stomatal size-density scaling relationship can be explained in terms of preserving a constant \gsmax{} that is proportional to $D S ^ {0.5}$ when $bm$ is constant (Equation \ref{eq:gsmax}). In other words, there are infinitely many combinations of $D$ and $S ^ {0.5}$ with the same \gsmax. If \gsmax{} is held constant at $C_g$, then the resulting size-density scaling relationship on a log-log scale is:

$$ d = c_g - 0.5 s $$
where lowercase variables are log-transformed equivalents of their uppercase counterparts. The scaling exponent $\beta_g = 0.5$ preseves $C_g$.

Next, imagine there is similarly a scaling exponent $\beta_p$ that preserves $p_\text{colonize}$ for the product $D S^{\beta_p}$. If $\beta_p = 0.5$, then $p_\text{colonize}$ is always proportional to \gsmax. If $\beta_p > 0.5$, small, densely packed stomata would be better defended (lower $p_\text{colonize}$) compared to larger, sparsely spaced stomata with the same \gsmax. If $\beta_p < 0.5$, small, densely packed stomata would be less defended (higher $p_\text{colonize}$) compared to larger, sparsely spaced stomata with the same \gsmax. I refer to the three outcomes ($\beta_p = 0.5$, $\beta_p < 0.5$, and $\beta_p > 0.5$) as iso-, hypo-, and hyper-conductance, respectively. I was unable to solve analytically for $\beta_p$, so I numerically calculated isoclines of $p_\text{colonize}$ over the grid of $D$ and $S$ values described in the preceding subsection. I numerically calculated the scaling relationships at a constant $p_\text{colonize} \in \{0.025, 0.05, 0.1, 0.2, 0.4\}$ for $H \in \{0, 0.01, 0.1\}$.

# Results {-}

I analyzed an idealized, spatially explicit [Model] of how a pathogen lands on a leaf and finds a stomate to colonize the leaf using a random search. To my knowledge, this is the first model that makes quantitative predictions about the relationship between stomatal anatomy, the probability of colonization, and their impact on stomatal size-density scaling.

## Nonlinear relationships between colonization, stomatal cover, and conductance {-}

The probability of colonization ($p_\text{colonize}$) is not simply a one-to-one relationship between the fraction of epidermal area allocated to stomata (\fs). At low \fs, $p_\text{colonize}$ increases faster rapdily relative to \fs{} at first Figure \ref{fig:fig3}a). At higher \fs, the $p_\text{colonize}$ increases linearly with \fs. When $H = 0$, any combination of stomatal size ($S$) and density ($D$) with the same \fs{} have the same effect on $p_\text{colonize}$. When $H > 0$, pathogens are less likely to land close enough to a stomate to infect before dying, so $p_\text{colonize}$ is closer to \fs{} (Figure \ref{fig:fig3}a). Furthermore, $p_\text{colonize}$ depends on $D$ and $S$, not just \fs. For the same \fs, leaves with greater $D$ have higher $p_\text{colonize}$ (Figure \ref{fig:fig3}a). Holding \fs{} constant, leaves with lower $D$ and higher $S$ will have a greater distance ($v_i$) between a pathogen and its stomata. When $H > 0$, this extra distance leads more pathogens to die before they can find a stomate. In contrast to \fs, $p_\text{colonize}$ increases at a greater than linear rate with \gsmax. Greater $D$ (smaller $S$) is associated with lower $p_\text{colonize}$ when \gsmax{} is held constant (Figure \ref{fig:fig3}b). This happens because $p_\text{colonize}$ increases approximately linearly with $S$ whereas \gsmax{} is proportional to $S^{0.5}$.

\begin{figure}
  \centering
	\includegraphics{../figures/fig3.pdf}
	\caption{\textbf{The probability of colonization increases with both stomatal cover and conductance} \textbf{a.} The probability of colonization ($p_\text{colonize}$, $y$-axis) initially increases rapidly with stomatal cover (\fs), then slows down to a linear relationship. Overall, $p_\text{colonize}$ is lower when pathogens can die on the leaf surface ($H > 0$). The relationship between \fs{} and $p_\text{colonize}$ is the same regardless of stomatal density when $H = 0$ (upper facet). When $H > 0$, higher density (solid lines) increase $p_\text{colonize}$ (lower facets). \textbf{b.} $p_\text{colonize}$ increases expoentially with \gsmax{} at all stomatal densities, but $p_\text{colonize}$ is much lower at higher densities for a given \gsmax. The relationship between \gsmax{} and $p_\text{colonize}$ is similar for all values of $H$.}
	\label{fig:fig3}
\end{figure}

## Hyper-conductance size-density scaling {-}

The scaling relationship between $S$ and $D$ that preserves $p_\text{colonize}$ is always greater 0.5 (hyper-conductance), but usually less than 1. When $H = 0$, the scaling relationship is essentially 1 (Figure \ref{fig:fig4}), which means that an increase \fs{} leads to a proportional increase in $p_\text{colonize}$. Because the scaling relationship is greater than 0.5, leaves with greater stomatal density will have lower $p_\text{colonize}$ than leaves lower stomatal density but the same \gsmax. In other words, increasing $D$ and lowering $S$ allows plants to reduce $p_\text{colonize}$ while maintaining \gsmax. The scaling relationship is slightly less than 1, but still greater than 0.5, when $H > 0$ (Figure \ref{fig:fig4}). In this area of parameter space, lower stomatal density can reduce \fs{} while $p_\text{colonize}$ is constant, but this will still result in lower \gsmax.

\begin{figure}
  \centering
	\includegraphics{../figures/fig4.pdf}
	\caption{Log-log scaling relationships between stomatal density ($D$, \textit{x}-axis) and size ($S$, \textit{y}-axis) that preserve the probability of colonization ($p_\text{colonize}$). In each panel, solid lines indicate values of $D$ and $S$ where $p_\text{colonize}$ is 0.025 (lowest line), 0.05, 0.1, 0.2, or 0.4 (highest line). For reference, dashed grey lines show scaling relationships that preserve \fs{} ($\beta = 1$, $\text{slope} = - 1 / \beta = -1$) and \gsmax{} ($\beta = 0.5$, $\text{slope} = - 1 / \beta = -2$) drawn through the centroid of the plotting region. When the death rate on the leaf surface is low ($H = 0$), the scaling exponent is very close to $\beta = 1$. When $H > 0$, $0.5 < \beta < 1$ and is slightly nonlinear on a log-log scale.}
	\label{fig:fig4}
\end{figure}

# Discussion {-}

Stomatal density and size set the upper limit on gas exchange in leaves [@harrison_influence_2019] and is often closely related to operational stomatal conductance in nature [@murray_consistent_2019].  Despite the fact that many ecologically and economically significant plant pathogens infect through stomata, the relationship between stomatal anatomy and susceptibility to foliar pathogens is less clear than it is for gas exchange. To develop testable predictions, we need mathematical models that can clarify the potential for tradeoffs between stomatal conductance, stomatal cover, and disease resistance. I used a spatially explicit model of a pathogen searching for a stomate to colonize a host. From this [Model], I derived predictions about the relationship between stomatal anatomy and disease resistance for the first time. The model predicts that the probability of colonization is not always proportional to the surface area of leaf covered by stomata (\fs), as one might intuitively predict. If the leaf surface is a hostile environment and pathogens have a limited time to search, lower stomatal density decreases the probability of colonization even if \fs{} is constant. However, \gsmax{} decreases proportionally more than the probability of colonization. The model reveals the potential for conflicting demands of maximizing disease resistance, minimizing stomatal cover, and maintaining stomatal conductance. Including the effect of anatomy on disease resistance therefore has the potential to change our understading of how stomatal size-density scaling evolves in land plants.

<!-- paragraph on first main result -->
The model predicts that in most cases, increasing stomatal cover should lead to a proportonal increase in susceptibility, which is the implicit assumption of some empirical studies (e.g. @mckown_association_2014; @tateda_host_2019; @dutton_bacterial_2019; @fetter_trade-offs_2019). It also makes new, testable predictions that are less intuitive. At very low \fs, there is a rapid increase in susceptibility (Figure \ref{fig:fig3}a). If there are no stomata, the probability of colonization is 0, so the first few stomata dramatically increase the probability. This is unlikely to be significant for abaxial (lower) leaf surfaces, which usually have most of the stomata [@salisbury_causes_1928; @metcalfe_anatomy_1950; @mott_adaptive_1984; @peat_comparative_1994; @jordan_using_2014; @muir_making_2015; @bucher_stomatal_2017; @drake_two_2019]. However, many adaxial (upper) leaf surfaces have zero or very few stomata. Using adaxial leaf surfaces, it should be possible to test if small changes in stomatal size or density have a larger effect on disease susceptibility when \fs{} is low. The nonlinear increase in $p_\text{colonize}$ is less apparent when $H > 0$ (Figure \ref{fig:fig3}a). A more hostile microenvironment (e.g. drier, higher UV) should therefore reduce the effect of increased size or density as low \fs. If true, the dimishing marginal effect of \fs{} on colonization could explain why stomatal ratio on the upper and lower surface is bimodal [@muir_making_2015]. The initial cost of adaxial (upper) stomata is high, but if the benefits outweigh the costs, then equal stomatal densities on each surface maximize CO$_2$ supply for photosynthesis [@parkhurst_adaptive_1978; @gutschick_photosynthesis_1984; @parkhurst_intercellular_1990].

<!-- paragraph on scaling -->
An effect of stomatal size and density on susceptibility to foliar pathogens could change our understanding of stomatal size-density scaling. Since allocating leaf epidermis to stomata may be costly [@franks_mechanical_2007; @assmann_guard_1987; @dow_physiological_2014; @lehmann_effects_2015; @baresch_competition_2019], selection should favor leaves that achieve a desired \gsmax{} while minimizing \fs{} [@de_boer_optimal_2016]. Because of their different scaling exponents (Equation \ref{eq:gsmax}, \ref{eq:fs}), smaller, densely packed stomata can achieve the same \gsmax{} at minimum \fs. However, many leaves have larger, sparsely packed stomata. Incorporating susceptibility to disease may explain why. If pathogens have a limited time to find stomata before dying ($H > 0$), then the scaling exponent between size and density that keeps $p_\text{colonize}$ constant is between 0.5 and 1, the scaling exponents for \gsmax{} and \fs, respectively (Figure \ref{fig:fig4}). Greater density of smaller stomata can increase \gsmax{} while keeping $p_\text{colonize}$ constant, but this will increase \fs. Conversely, \fs{} could decrease while keeping $p_\text{colonize}$ constant, but this will decrease \gsmax. This sets up the potential for conflict between competing goals. The optimal stomatal size and density will therefore depend on the precise costs and benefits of infection, stomatal conductance, and stomatal cover. This may explain why many leaves have large, sparsely packed stomata despite the fact that they could achieve the same \gsmax{} and lower \fs{} with smaller, more densely packed stomata.

<!-- scope of model -->
The model examines the probability of colonization for a single pathogen. The calculated probabilities of colonization should not be interpreted as exact predictions, but rather as depicting qualitative relationships between stomatal anatomy and infection severity. The model is most applicable to diseases where the host has some resistance. The energetic cost and lost photosynthetic capacity (closed stomata, necrosis, etc.) of dealing with a pathogen is assumed to be proportional to the amount of infection. The actual fitness cost will be modulated by the number of pathogens landing on the leaf and the cost of infection. In environments with fewer or less virulent pathogens, the fitness cost of infection will be less than in environments with more abundant, virulent pathogens.
The model is less relevant to very susceptible host plants that can be severely damaged or killed by a small number of colonizations that spread unchecked throughout the host tissue.

<!-- other limitations -->
The purpose of this model is to provide a general foundation to examine the relationship between stomatal size, density, and defense against foliar pathogens. In its generality, it overlooks interesting natural history and biologically important features of specific plants and their pathogens. For example, some pathogens actively seek out and find stomata [@kiefer_host_2002], whereas I assumed that pathogens randomly orient themselves on the leaf. Including sensing should increase the probability of colonization. I also assumed that the plant does not respond by, for example, closing stomata when it senses a pathogen. However, plants can sense and close stomata when pathogens land on the leaf, but pathogens can pry stomata back open [@melotto_plant_2006]. Future work on specific plant-pathogen interactions could build on this model by adding more biological realism to provide more precise predictions.

# Conclusion {-}

The model makes two non-intuitive predictions. First, the effect of increased stomatal density or size on susceptibility to foliar pathogens is greatest when stomatal cover is very low. Second, maximizing disease resistance sets up a potential conflict between minimizing stomatal cover and maximizing stomatal conductance. The first prediction may be relatively straightforward to test experimentally with adaxial (upper) stomata that occur at low and moderate densities within the same or closely related species [@muir_morphological_2014; @mckown_association_2014; @fetter_trade-offs_2019]. The second prediction about size-density scaling is more complex because we would need to know the relationships between colonization, stomatal cover, stomatal conductance, and fitness in natural conditions. Testing these predictions in a variety of species would help determine whether pathogens have played an important role shaping  stomatal anatomy in land plants.

<!-- # Acknowledgements {-} -->

# Funding {-}

I am grateful startup funds from the University of Hawaii for supporting this work.

# Conflict of Interest Statement {-}

The author declares that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.

# Supplementary Material {-}

I calculated \gsmax{} (Equation \ref{eq:gsmax}) to water vapor at a reference leaf temperature ($T_\mathrm{leaf} = 25 ^ \circ$ C) following @sack_developmental_2016. They defined a biophysical and morphological constant as:

\begin{align*}
  b = & D_\text{wv} / v \\
  m = & \frac{\pi c ^ 2}{j ^ {0.5} (4 h j + \pi c)}
\end{align*}

$b$ is the diffusion coefficient of water vapor in air ($D_\text{wv}$) divided by the kinematic viscosity of dry air ($v$). $D_\text{wv} = 2.49 \times 10^{-5} ~ \text{m}^2 ~ \text{s}^{-1}$ and $v = 2.24 \times 10 ^ {-2} ~ \text{m}^3 ~ \text{mol} ^ {-1}$ at $25 ^ \circ$ [@monteith_principles_2013]. For kidney-shaped guard cells, $c = h = j = 0.5$.

# References {-}