### 2019-12-10 ms/ms.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is just an example/guide for you to refer to when submitting manuscripts to Frontiers, it is not mandatory to use Frontiers .cls files nor frontiers.tex  %
% This will only generate the Manuscript, the final article will be typeset by Frontiers after acceptance.
%                                              %
%                                                                                                                                                         %
% When submitting your files, remember to upload this *tex file, the pdf generated with it, the *bib file (if bibliography is not within the *tex) and all the figures.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Version 3.4 Generated 2018/06/15 %%%
%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%

\documentclass[utf8]{frontiersSCNS}

%\setcitestyle{square} % for Physics and Applied Mathematics and Statistics articles
\usepackage{url,hyperref,lineno,microtype,subcaption}
\usepackage[onehalfspacing]{setspace}

\linenumbers

% BELOW TAKEN FROM rticles plos template
%
% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% hyperref package, useful for hyperlinks
\usepackage{hyperref}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% Sweave(-like)
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
\newenvironment{Schunk}{}{}
\DefineVerbatimEnvironment{Code}{Verbatim}{}
\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
\newenvironment{CodeChunk}{}{}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Below is from frontiers
%
\bibliographystyle{frontiersinSCNS}
% Use doublespacing - comment out for single spacing
%\usepackage{setspace}
%\doublespacing

% Leave a blank line between paragraphs instead of using \\

\def\keyFont{\fontsize{8}{11}\helveticabold }

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW

%% END MACROS SECTION

\usepackage{longtable}
\usepackage{tikz}

% short-cut for fs, gmax, and gop
\newcommand{\fs}{$f_\text{S}$}
\newcommand{\gsmax}{$g_\text{s,max}$}
\newcommand{\gsop}{$g_\text{s,op}$}

\def\Authors{
  Christopher D. Muir\,\textsuperscript{1*}}

\def\Address{

  \textsuperscript{1} School of Life Sciences, University of Hawaii,  Honolulu,  Hawaii,  USA
  }

  \def\corrAuthor{Christopher D. Muir}\def\corrAddress{,  }\def\corrEmail{\href{mailto:cdmuir@hawaii.edu}{\nolinkurl{cdmuir@hawaii.edu}}}
  \def\firstAuthorLast{Muir {et~al.}}

\begin{document}
\onecolumn
\firstpage{1}

\title[Stomata tradeoff photosynthesis for defense]{A stomatal model of anatomical tradeoffs between photosynthesis and
pathogen defense}
\author[\firstAuthorLast]{\Authors}
\address{} %This field will be automatically populated
\correspondance{} %This field will be automatically populated

\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}

\maketitle

\begin{abstract}

Stomatal pores control both leaf gas exchange and are an entry for many plant pathogens, setting up the potential for tradeoffs between photosynthesis and defense. To prevent colonization and limit infection, plants close their stomata after recognizing pathogens. In addition to closing stomata, anatmoical shifts to lower stomatal density and/or size may also limit pathogen colonization, but such developmental changes would permanently reduce the gas exchange capacity for the life of the leaf. I developed and analyzed a spatially explicit model of pathogen colonization on the leaf as a function of stomatal size and density, anatomical traits which determine maximum rates of gas exchange. The model predicts greater stomatal size or density increases colonization, but the effect is most pronounced when stomatal cover is low. I also derived scaling relationships between stomatal size and density that preserves a given probability of colonization. These scaling relationships set up a potential conflict between maximizing defense and minimizing stomatal cover. To my knowledge, this is the first mathematical model connecting gas exchange and pathogen defense via stomatal anatomy. It makes predictions that can be tested with experiments and may explain variation in stomatal anatomy among plants. The model is generalizable to many types of pathogens, but lacks significant biological realism that may be needed for accurate predictions.

\tiny
 \keyFont{ \section{Keywords:} anatomy, leaf gas exchange, model, pathogen, photosynthesis, scaling, stomata, tradeoff }

\end{abstract}

\hypertarget{introduction}{%
\section*{Introduction}\label{introduction}}
\addcontentsline{toc}{section}{Introduction}

Stomata evolved to regulate gas exchange in and out of the leaf
(Hetherington and Woodward, 2003; Berry et al., 2010; Chater et al.,
2017), but many plant pathogens take advantage of these chinks in the
leaf cuticular armor to infect prospective hosts (Zeng et al., 2010;
McLachlan et al., 2014; Melotto et al., 2017). The density and size of
stomata set the anatomical maximum rate of stomatal conductance to
CO\(_2\) and water vapor (Brown and Escombe, 1900; Parlange and
Waggoner, 1970; Franks and Farquhar, 2001; Franks and Beerling, 2009b;
Lehmann and Or, 2015; Sack and Buckley, 2016; Harrison et al., 2019),
but the pore area shrinks and expands in response to internal and
external factors to regulate gas exchange dynamically (Buckley, 2019).
Many plant pathogens, including viruses (Murray et al., 2016), bacteria
(Melotto et al., 2006; Underwood et al., 2007), protists (Fawke et al.,
2015), and fungi (Hoch et al., 1987; Zeng et al., 2010) use stomatal
pores to gain entry into the leaf. Since stomatal conductance is a major
limitation on photosynthesis (Farquhar and Sharkey, 1982; Jones, 1985)
while pathogens reduce fitness, this sets up a potential tradeoff
between increased photosynthesis and defense against pathogens. Although
there have been many empirical studies on the effect that pathogens have
on stomata, there is no theoretical framework in which to place these
findings. Lack of a theoreatical framework makes it difficult to answer
general questions about how selection for pathogen defense constrains
maximum rates of gas exchange.

Stomatal anatomy is the key link between gas exchange and pathogen
colonization. The density and size of stomata not only determines the
theoretical maximum stomatal conductance (\gsmax), but is also
proportional to the operational stomatal conductance (\gsop) in many
circumstances (Franks et al., 2009, 2014; Dow et al., 2014a; McElwain et
al., 2016; Murray et al., 2019). Therefore, I use anatomical \gsmax{} as
a proxy for \gsop{} and do not address dynamic changes in stomatal
aperture (see \protect\hyperlink{discussion}{Discussion}). Harrison et
al. (2019) recently reviewed the relationship between stomatal anatomy
and gas exchange in detail.

Many pathogens rely on stomata to gain entry into the leaf, but it is
unclear how anatomical traits that determine \gsmax{} (size and density)
affect the ability of pathogens to colonize leaves. The impact of
pathogens on host fitness is complex, but after a pathogen reaches a
host, the first major step is colonization. Once infected, a pathogen
can reduce fitness. Susceptible hosts can lose much of their biomass or
die, but even resistant hosts must allocate resources to defense or
reduce photosynthesis because of necrosis around sites of infection.
Plants can limit colonization physiologically by closing stomata after
they recognize pathogens, called stomatal defense (Melotto et al.,
2017). Anatomy may be another layer of defense; plants can reduce
pathogen colonization by developing leaves with lower stomatal density
and/or size. Infection increases in leaves with higher stomatal density
(McKown et al., 2014; Tateda et al., 2019; Dutton et al., 2019; Fetter
et al., 2019), which suggests that both anatomical and physiological
responses (stomatal closure) affect host colonization. By the same
logic, if stomatal density were held constant, larger stomata should
also increase colonization because they occupy more area. One key
difference between physiological and anatomical defenses is that
stomatal closure reduces gas exchange transiently, whereas anatomy
constrains gas exchange throught the life of the leaf. Because we do not
understand the relationship between stomatal anatomy and colonization
well, we cannot predict the relationship between \gsmax{} and infection.

If stomatal size and density affect pathogen colonization, selection to
limit colonization may shape stomatal size-density scaling
relationships. Botanists have long recognized that stomatal size and
density are inversely correlated (Weiss, 1865; Tichá, 1982; Hetherington
and Woodward, 2003; Sack et al., 2003; Franks and Beerling, 2009a;
Brodribb et al., 2013; de Boer et al., 2016), but the evolutionary
origin of this relationship is not yet known. Here I argue that
pathogens could shape selection on this relationship. Explanations for
inverse size-density scaling are usually cast in terms of preserving
\gsmax{} and/or total epidermal area allocated to stomata (\fs; de Boer
et al., 2016) because there are many combinations of stomatal size and
density that have same \gsmax{} or same \fs{}:

\begin{align}
  g_\text{s,max} & = b m D S ^ {0.5} \label{eq:gsmax} \\
  f_\text{S} & = D S \label{eq:fs}.
\end{align}

\(b\) and \(m\) are assumed to be biophysical and morphological
constants, \emph{sensu} Sack and Buckley (2016) (see
\protect\hyperlink{supplementary-material}{Supplementary Material}). If
size and density also affect pathogen colonization, then selection from
foliar pathogens could significantly alter the size-density scaling
relationship. The empirical size-density scaling relationship is linear
on a log-log scale, determined by an intercept \(\alpha\) and slope
\(\beta\):

\begin{align}
  D & = e ^ \alpha S ^ {-\beta}; \\
  d & = \alpha - \beta S.
\end{align}

For brevity, \(d = \text{log}(D)\) and \(s = \text{log}(S)\).
Rearranging Equations \ref{eq:gsmax} and \ref{eq:fs}, a scaling
relationship where \(\beta = 0.5\) preserves \gsmax{} while
\(\beta = 1\) preserves \fs.

How would adding pathogens alter these predicted scaling relationships?
For simplicity, imagine two environments, one without foliar pathogens
and one with lots. In the absence of foliar pathogens, we expect
size-density scaling to preserve \gsmax{}, \fs{}, or some least-cost
combination of them. What happens when we introduce pathogens? Assuming
that stomatal size and density increase pathogen colonization, then
selection will favor reduced size and/or density. This would change the
intercept \(\alpha\) but not the slope. The effect of foliar pathogens
on the slope depends on the relationship between size, density, and
probability of colonization. If the probability of colonization is
proportional to the product of \emph{linear} stomatal size
(\(S ^ {0.5}\)) and density (\(\propto D S ^ {0.5}\) as for \gsmax) then
it has the same effect on the slope as \gsmax{} because there are many
combinations of \(D\) and \(S ^ {0.5}\) that have same probability of
colonization. If the probability of colonization is proportional to the
product of \emph{areal} stomatal size (\(S\)) and density
(\(\propto D S\) as for \fs) then it has the same effect on the slope as
\fs{} because there are many combinations of \(D\) and \(S\) that have
same probability of colonization. Alternatively, the probability of
colonization may have a different scaling relationsip (neither 0.5 nor
1) or may be nonlinear on a log-log scale. Unlike \gsmax{} and \fs{}, we
do not have theory to predict a stomatal size-density relationship that
preserves the probability of colonization.

In summary, the physical relationship between stomatal size, density,
and conductance is well established (Harrison et al., 2019). The same
traits likely affect the probability of pathogen colonization, but we do
not have a theoretical model that makes quantitative predictions. The
inverse stomatal size-density relationship has usually been explained in
terms of preserving stomatal conductance and/or stomatal cover, but
selection by pathogens might alter scaling. To address these gaps, the
goals of this study are to 1) introduce a spatially explicit model
pathogen colonization on the leaf surface; 2) use the model to predict
the relationship between \gsmax, \fs{}, and the probability of
colonization; 3) work out what these relationships predict about
stomatal size-density scaling.

\hypertarget{model}{%
\section*{Model}\label{model}}
\addcontentsline{toc}{section}{Model}

In this section, I introduce a spatially explicit model of pathogen
colonization on a leaf surface. I explain the model structure and
assumptions here; the
\protect\hyperlink{materials-and-methods}{Materials and Methods} section
below describes how I analyzed the model to address the goals of the
study. For generality, I refer to a generic ``pathogen'' that lands on
leaf and moves to a stomate. The model is agnostic to the type of
pathogen (virus, bacterium, fungus, etc.) and the specific biological
details of how it moves (biotrophy). For example, motile bacterial cells
can land and move around where fungi may germinate from a cyst and grow
until they form an appresorium for infection. These very different
biotropic movements on the leaf are treated identically here. I used
Sympy version 1.3 (Meurer et al., 2017) for symbolic derivations.

\textbf{Table 1} \textbar{} Glossary of mathematical symbols. The
columns indicate the mathematical Symbol used in the paper, the
associated symbol used in R scripts, scientific Units, and a verbal
Description.

\begin{longtable}[]{@{}llll@{}}
\toprule
\begin{minipage}[b]{0.09\columnwidth}\raggedright
Symbol\strut
\end{minipage} & \begin{minipage}[b]{0.09\columnwidth}\raggedright
R\strut
\end{minipage} & \begin{minipage}[b]{0.09\columnwidth}\raggedright
Units\strut
\end{minipage} & \begin{minipage}[b]{0.61\columnwidth}\raggedright
Description\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.09\columnwidth}\raggedright
\(D,d\)\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright
\texttt{D,\ d}\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright
\(\textrm{mm}^{-2}\)\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright
stomatal density (\(d = \text{log}~D\))\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright
\(f_s\)\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright
\texttt{f\_s}\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright
none\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright
stomatal cover (\(f_s = D S\))\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright
\(g_\text{s,max}\)\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright
\texttt{g\_smax}\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright
\(\text{mol}~\text{m}^{-2}~\text{s}^{-1}\)\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright
anatomical maximum stomatal conductance\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright
\(g_\text{s,op}\)\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright
\texttt{g\_sop}\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright
\(\text{mol}~\text{m}^{-2}~\text{s}^{-1}\)\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright
operational stomatal conductance\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright
\(H\)\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright
\texttt{H}\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright
\(\mu\textrm{m} ^{-1}\)\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright
death rate of pathogen on leaf surface\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright
\(R\)\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright
\texttt{R}\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright
\(\mu\textrm{m}\)\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright
stomatal radius (\(S = 2 \pi R ^ 2\))\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright
\(S,s\)\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright
\texttt{S,\ s}\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright
\(\mu\textrm{m}^2\)\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright
stomatal size (\(s = \text{log}~S\))\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright
\(\theta_i\)\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright
\texttt{theta\_i}\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright
radians\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright
angles between pathogen (\(x_p,y_p\)) and lines tangent to the
circumfrence of stomate \(i\)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright
\(U\)\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright
\texttt{U}\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright
\(\mu\textrm{m}\)\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright
interstomatal distance\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright
\(v_i\)\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright
\texttt{v\_i}\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright
\(\mu\textrm{m}\)\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright
distance between pathogen (\(x_p,y_p\)) and stomate \(i\)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright
\(x_i,y_i\)\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright
\texttt{x\_i,y\_i}\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright
\(\mu\textrm{m}\)\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright
position of stomate \(i\)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright
\(x_p,y_p\)\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright
\texttt{x\_p,y\_p}\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright
\(\mu\textrm{m}\)\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright
starting position of pathogen\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\hypertarget{spatial-representation-of-stomata}{%
\subsection*{Spatial representation of
stomata}\label{spatial-representation-of-stomata}}
\addcontentsline{toc}{subsection}{Spatial representation of stomata}

Stomata develop relatively equal spacing to minimize resistance to
lateral diffusion (Morison et al., 2005), allow space between stomata
(Dow et al., 2014b), and prevent stomatal interference (Lehmann and Or,
2015). Here I assume that stomata are arrayed in an equilateral
triangular grid with a density \(D\) and size (area) \(S\). This
assumption ignores veins, trichomes, and within-leaf variation in
stomatal density. Stomata are therefore arrayed in an evenly spaced grid
(Figure \ref{fig:fig1}a). The interstomatal distance \(U\), measured as
the distance from the center of one stomata to the next, is the maximal
diagonal of the hexagon in \(\mu\textrm{m}\) that forms an equal area
boundary between neighbording stomata. The area of a hexagon is
\(A_\textrm{hexagon} = \frac{\sqrt{3}}{2} U ^ 2\). By definition the
stomatal density is the inverse of this area, such that
\(D = A^{-1}_\textrm{hexagon} = \frac{2}{\sqrt{3}} U ^ {-2}\).
Therefore, interstomatal distance can be derived from the stomatal
density as:

\[ D = \frac{2}{\sqrt{3}} U ^{-2} \]
\[ U = \bigg(\frac{2}{\sqrt{3}} D ^{-1}\bigg) ^ {0.5} \]

For example, if the density is
\(D = 10 ^ 2~\textrm{mm}^{-2} = 10 ^ {-4}~\mu\textrm{m}^{-2}\), then
\(U\) is 107.5 \(\mu \textrm{m}\). Parkhurst (1994) described this
result previously. I also make the simplifying assumption that stomata
are perfectly circular with radius \(R\). This may be approximately true
for fully open stomata with kidney-shaped guard cells. Although I assume
stomata are circular here, in calculating \gsmax, I assume typical
allometric relationships between length, width, and pore area (see
\protect\hyperlink{supplementary-material}{Supplementary Material}).

\begin{figure}
  \centering
    \includegraphics[width=6.5in]{../figures/fig1.jpeg}
    \caption{\textbf{A spatially explicit model of stomatal anatomy and pathogen colonization}. \textbf{a}. Stomata are assumed to be in a homogenous equilateral triangular grid, which means that we can extrapolate from \textbf{b.} a focal triangle to the entire leaf. The circles represent idealized stomata; the grey lines between them are for visualization. \textbf{c.} By symmetry, a single focal region within the focal triangle can be modeled and extrapolated to the rest of the area. \textbf{d.} The model assumes that a pathogen, depicted as a grey rod, lands somewhere on the leaf surface and will sucessfully locate a stomate if it moves at the correct angle, depicted by the grey polygons.}
    \label{fig:fig1}
\end{figure}

\hypertarget{spatial-representation-of-pathogen-search}{%
\subsection*{Spatial representation of pathogen
search}\label{spatial-representation-of-pathogen-search}}
\addcontentsline{toc}{subsection}{Spatial representation of pathogen
search}

Now imagine that a pathogen lands at a uniform random position within
the focal region and must arrive at a stomate to colonize. If it lands
on a stomate, then it infects the leaf with probability 1; if it lands
between stomata, then it infects the leaf with probability
\(p_\text{locate}\). This is the probability that it locates a stomate,
which I will derive below. The probabilities of landing on or between a
stomate are \fs{} and \(1 - f_S\), respectively. Hence, the total
probability of colonization is:

\begin{equation}
  \label{eq:p_colonize}
  p_\text{colonize} = f_\text{S} + (1 - f_\text{S}) p_\text{locate}.
\end{equation}

I assume that the pathogen cannot sense where stomata are and orients at
random, thereafter traveling in that direction. If it successfully
locates a stomate, it colonizes the leaf, but otherwise does not infect.
If there is a high density of stomata and/or large stomata, the
probability of locating a stomate increases. By assuming that stomata
form an equilateral triangular grid (see above), we can extrapolate what
happens in a focal triangle (Figure \ref{fig:fig1}b) by symmetry.
Further, since an equilateral triangle can be broken up into six
identical units (Figure \ref{fig:fig1}c), we can simply calculate
\(p_\text{locate}\) in this focal area. This implicitly assumes that the
probability of colonzing stomata outside the focal area is 0 because
they are too far away.

Imagine that the pathogens lands in position (\(x_p,y_p\)) within the
triangle. The centroid of the triangle is at position (\(x_c,y_c\)) and
a reference stomate is at position \((0,0)\) (Figure \ref{fig:fig2}a).
Therefore \(x_c = U / 2\) and \(y_c = \sqrt{3} U / 6\). The other
stomata are at positions \((U/2, \sqrt{3} U /2)\) and \((U,0)\) (Figure
\ref{fig:fig2}). \(x_p\) and \(y_p\) are defined as the horizontal and
vertical distances, respectively, from the pathogen to the reference
stomate at position \((0,0)\).

Given that the pathogen starts at position (\(x_p,y_p\)), what's the
probability of contacting one of the stomata at the vertices of the
focal triangle? I assume the probability of contacting a stomate is
equal to the proportion of angular directions that lead to a stomate
(Figure \ref{fig:fig1}d). I solved this by finding the angles
(\(\theta_1, \theta_2, \theta_3\)) between lines that are tangent to the
outside of the three stomata and pass through (\(x_p,y_p\)) (Figure
\ref{fig:fig2}a). If stomate \(i\) is centered at (\(x_i,y_i\)), the two
slopes of tangency as function of pathogen position are:

\begin{align}
  t_{i,1}(x_p,y_p) & = \frac{-R e_{i,2}(x_p,y_p) + e_{i,3}(x_p,y_p)}{e_{i,1}(x_p,y_p)} \\
  t_{i,2}(x_p,y_p) & = \frac{R e_{i,2}(x_p,y_p) + e_{i,3}(x_p,y_p)}{e_{i,1}(x_p,y_p)}
\end{align}

where

\begin{align}
  e_{i,1}(x_p,y_p) & = (R^2 - x_i^2 + 2 x_i x_p - x_p^2), \\
  e_{i,2}(x_p,y_p) & = \sqrt{-e_{i,1} + (y_i - y_p) ^ 2}, \\
  e_{i,3}(x_p,y_p) & = - x_i y_i + x_i y_p + x_p y_i - x_p y_p.
\end{align}

Note that \(i \in \{1, 2, 3\}\), indexing the three stomata in the focal
triangle. The angle in radians between \(t_{i,1}(x_p,y_p)\) and
\(t_{i,2}(x_p,y_p)\) is:

\begin{equation}
  \theta_i(x_p,y_p) = \text{arctan}\bigg(\frac{t_{i,1}(x_p,y_p) - t_{i,2}(x_p,y_p)}{1 + (t_{i,1}(x_p,y_p) t_{i,2}(x_p,y_p))}\bigg)
\end{equation}

\begin{figure}
\centering
\input{../figures/fig2.tex}
\caption{\textbf{Spatial representation of stomata and pathogen.} \textbf{a.} The pathogen starts at a uniform random position within the focal region denoted ($x_p,y_p$). Within the focal triangle, the reference stomata is at position (0,0) by definition, and other stomatal positions are determined by the interstomatal distance $U$. \textbf{b.} Within the focal region, a pathogen can land within the stomate (white cirlce with grey outline and radius $R$) or in the grey area. The outer borders of this area are shown and depend on $R$ and $U$. For a given position $x$, there is a minimum $y$-value ($y_\text{min}$, dashed line) and maximum $y$-value ($y_\text{max}$, solid line).}
\label{fig:fig2}
\end{figure}

I further assumed that the longer distance a pathogen must travel, the
less likely it would be to locate a stomate. For example, if stomata are
at very low density, then a pathogen may die before it reaches a stomate
because of UV, dessication, or another factor. I included this effect by
assuming the probability of reaching a stomate declines exponentially at
rate \(H\) with the Euclidean distance \(v_i(x_p,y_p)\) between the
pathogen location and the edge of stomata \(i\), which is distance \(R\)
from its center at \(x_i,y_i\):

\begin{equation}
  v_i(x_p,y_p) = \sqrt{(x_i - x_p) ^ 2 + (y_i - y_p) ^ 2} - R.
\end{equation}

The probability of locating a stomate as a function of \(x_p\) and
\(y_p\) (\(f_\text{locate}(x_p,y_p)\)) is the sum of the angles divided
by \(2\pi\), discounted by their distance from the stomate:

\begin{equation}
  f_\text{locate}(x_p,y_p) = \frac{1}{2 \pi} \sum_{i=1}^3 e^{-H v_i(x_p,y_p)} \theta_i(x_p,y_p)
\end{equation}

When \(H = 0\), \(p_\text{locate}\) is the fraction of angles that lead
from \((x_p,y_p)\) to a stomate. When \(H > 0\), \(p_\text{locate}\) is
proportional to this fraction, but less than it depending on stomatal
density, size, and starting location of the pathogen.

To obtain the average \(p_\text{locate}\), we must integrate
\(f_\text{locate}(x_p,y_p)\) over all possible starting positions
\((x_p,y_p)\) within the focal area. The focal area is a 30-60-90
triangle with vertices at the center of the reference stomate \((0,0)\),
the midpoint of baseline \((U/2,0)\), and the centroid of the focal
trinagle \((U/2, \sqrt{3} / 6 U)\) (Figure \ref{fig:fig1}c).
Colonization occurs with probability 1 if the pathogen lands in the
reference stomate, so we need to integrate the probability of
colonization if it lands elsewhere. This region extends from the edge of
the stomate, at \(\sqrt{3} / 2 R\) to \(U/2\) (Figure \ref{fig:fig2}b).
At any \(x\), we integrate from the bottom of the focal area
(\(y_\text{min}\)) to the top (\(y_\text{max}\)):

\begin{align}
  y_\text{min} = f(x) & = \begin{cases}
    \sqrt{R ^ 2 - x ^ 2}, & \text{if } \frac{\sqrt{3}}{2} R < x < R \\ 
    0, & \text{if } R \leq x \leq \frac{U}{2} \end{cases} \\
  y_\text{max} = f(x) & = \frac{\sqrt{3}}{3} x
\end{align}

The integral is:

\begin{equation}
  p_\text{locate} = \frac{1}{a_\text{focal}}\int\limits_{\frac{\sqrt{3}}{2} R} ^ {U / 2} \int\limits_{y_\mathrm{min}} ^ {y_\mathrm{max}} f_\text{locate}(x, y) \,dx \,dy
\end{equation}

\(a_\text{focal}\) is the area of the focal region depicted in grey in
Figure \ref{fig:fig2}b:

\[ a_\text{focal} = \frac{U ^ 2}{8 \sqrt{3}} - \frac{\pi R ^ 2}{12} \]

\hypertarget{materials-and-methods}{%
\section*{Materials and Methods}\label{materials-and-methods}}
\addcontentsline{toc}{section}{Materials and Methods}

The \protect\hyperlink{model}{Model} calculates a probability of host
colonization (Equation \ref{eq:p_colonize}) as a function stomatal
density, size, and position of a pathogen on the leaf. I solved
\(p_\text{colonize}\) using the \texttt{integral2()} function in the
\textbf{pracma} package version 2.2.5 (Borchers, 2019) for numerical
integration. I used R version 3.6.1 (R Core Team, 2019) for all analyses
and wrote the paper in \textbf{rmarkdown} version 1.17 (Xie et al.,
2018; Allaire et al., 2019). Source code is deposited on GitHub
(\url{https://github.com/cdmuir/stomata-tradeoff}) and will be arcived
on Zenodo upon publication.

\hypertarget{what-is-the-relationship-between-stomatal-size-density-and-colonization}{%
\subsection*{What is the relationship between stomatal size, density,
and
colonization?}\label{what-is-the-relationship-between-stomatal-size-density-and-colonization}}
\addcontentsline{toc}{subsection}{What is the relationship between
stomatal size, density, and colonization?}

I calculated \(p_\text{colonize}\) over a biologically plausible grid of
stomatal size and density for hypostomatous species based on de Boer et
al. (2016). Stomatal density ranges from
\(10^1 - 10^{3.5}~\text{mm}^{-2}\); stomatal size ranges from
\(10^1 - 10^{3.5}~\mu\text{m}^2\). I only considered combinations of
size and density where \fs{} was less than 1/3. For simplicity, I have
not extended the current analysis to amphistomatous leaves. I crossed
stomatal traits with three levels of \(H \in \{0, 0.01, 0.1\}\). When
\(H = 0\), a pathogen perists indefinitely on the leaf surface.
\(H = 0.01\) and \(H = 0.1\) correspond to low and high death rates,
respectively. These values are not necessarily realistic, but illustrate
qualitatively how a hostile environment on the leaf surface alters model
predictions.

\hypertarget{how-do-pathogens-alter-optimal-stomatal-size-density-scaling}{%
\subsection*{How do pathogens alter optimal stomatal size-density
scaling?}\label{how-do-pathogens-alter-optimal-stomatal-size-density-scaling}}
\addcontentsline{toc}{subsection}{How do pathogens alter optimal
stomatal size-density scaling?}

The stomatal size-density scaling relationship can be explained in terms
of preserving a constant \gsmax{} that is proportional to
\(D S ^ {0.5}\) when \(bm\) is constant (Equation \ref{eq:gsmax}). In
other words, there are infinitely many combinations of \(D\) and
\(S ^ {0.5}\) with the same \gsmax. If \gsmax{} is held constant at
\(C_g\), then the resulting size-density scaling relationship on a
log-log scale is:

\[ d = c_g - 0.5 s \] where lowercase variables are log-transformed
equivalents of their uppercase counterparts. The scaling exponent
\(\beta_g = 0.5\) preseves \(C_g\).

Next, imagine there is similarly a scaling exponent \(\beta_p\) that
preserves \(p_\text{colonize}\) for the product \(D S^{\beta_p}\). If
\(\beta_p = 0.5\), then \(p_\text{colonize}\) is always proportional to
\gsmax. If \(\beta_p > 0.5\), small, densely packed stomata would be
better defended (lower \(p_\text{colonize}\)) compared to larger,
sparsely spaced stomata with the same \gsmax. If \(\beta_p < 0.5\),
small, densely packed stomata would be less defended (higher
\(p_\text{colonize}\)) compared to larger, sparsely spaced stomata with
the same \gsmax. I refer to the three outcomes (\(\beta_p = 0.5\),
\(\beta_p < 0.5\), and \(\beta_p > 0.5\)) as iso-, hypo-, and
hyper-conductance, respectively. I was unable to solve analytically for
\(\beta_p\), so I numerically calculated isoclines of
\(p_\text{colonize}\) over the grid of \(D\) and \(S\) values described
in the preceding subsection. I numerically calculated the scaling
relationships at a constant
\(p_\text{colonize} \in \{0.025, 0.05, 0.1, 0.2, 0.4\}\) for
\(H \in \{0, 0.01, 0.1\}\).

\hypertarget{results}{%
\section*{Results}\label{results}}
\addcontentsline{toc}{section}{Results}

I analyzed an idealized, spatially explicit
\protect\hyperlink{model}{Model} of how a pathogen lands on a leaf and
finds a stomate to colonize the leaf using a random search. To my
knowledge, this is the first model that makes quantitative predictions
about the relationship between stomatal anatomy, the probability of
colonization, and their impact on stomatal size-density scaling.

\hypertarget{nonlinear-relationships-between-colonization-stomatal-cover-and-conductance}{%
\subsection*{Nonlinear relationships between colonization, stomatal
cover, and
conductance}\label{nonlinear-relationships-between-colonization-stomatal-cover-and-conductance}}
\addcontentsline{toc}{subsection}{Nonlinear relationships between
colonization, stomatal cover, and conductance}

The probability of colonization (\(p_\text{colonize}\)) is not simply a
one-to-one relationship between the fraction of epidermal area allocated
to stomata (\fs). At low \fs, \(p_\text{colonize}\) increases faster
rapdily relative to \fs{} at first Figure \ref{fig:fig3}a). At higher
\fs, the \(p_\text{colonize}\) increases linearly with \fs. When
\(H = 0\), any combination of stomatal size (\(S\)) and density (\(D\))
with the same \fs{} have the same effect on \(p_\text{colonize}\). When
\(H > 0\), pathogens are less likely to land close enough to a stomate
to infect before dying, so \(p_\text{colonize}\) is closer to \fs{}
(Figure \ref{fig:fig3}a). Furthermore, \(p_\text{colonize}\) depends on
\(D\) and \(S\), not just \fs. For the same \fs, leaves with greater
\(D\) have higher \(p_\text{colonize}\) (Figure \ref{fig:fig3}a).
Holding \fs{} constant, leaves with lower \(D\) and higher \(S\) will
have a greater distance (\(v_i\)) between a pathogen and its stomata.
When \(H > 0\), this extra distance leads more pathogens to die before
they can find a stomate. In contrast to \fs, \(p_\text{colonize}\)
increases at a greater than linear rate with \gsmax. Greater \(D\)
(smaller \(S\)) is associated with lower \(p_\text{colonize}\) when
\gsmax{} is held constant (Figure \ref{fig:fig3}b). This happens because
\(p_\text{colonize}\) increases approximately linearly with \(S\)
whereas \gsmax{} is proportional to \(S^{0.5}\).

\begin{figure}
  \centering
    \includegraphics{../figures/fig3.pdf}
    \caption{\textbf{The probability of colonization increases with both stomatal cover and conductance} \textbf{a.} The probability of colonization ($p_\text{colonize}$, $y$-axis) initially increases rapidly with stomatal cover (\fs), then slows down to a linear relationship. Overall, $p_\text{colonize}$ is lower when pathogens can die on the leaf surface ($H > 0$). The relationship between \fs{} and $p_\text{colonize}$ is the same regardless of stomatal density when $H = 0$ (upper facet). When $H > 0$, higher density (solid lines) increase $p_\text{colonize}$ (lower facets). \textbf{b.} $p_\text{colonize}$ increases expoentially with \gsmax{} at all stomatal densities, but $p_\text{colonize}$ is much lower at higher densities for a given \gsmax. The relationship between \gsmax{} and $p_\text{colonize}$ is similar for all values of $H$.}
    \label{fig:fig3}
\end{figure}

\hypertarget{hyper-conductance-size-density-scaling}{%
\subsection*{Hyper-conductance size-density
scaling}\label{hyper-conductance-size-density-scaling}}
\addcontentsline{toc}{subsection}{Hyper-conductance size-density
scaling}

The scaling relationship between \(S\) and \(D\) that preserves
\(p_\text{colonize}\) is always greater 0.5 (hyper-conductance), but
usually less than 1. When \(H = 0\), the scaling relationship is
essentially 1 (Figure \ref{fig:fig4}), which means that an increase
\fs{} leads to a proportional increase in \(p_\text{colonize}\). Because
the scaling relationship is greater than 0.5, leaves with greater
stomatal density will have lower \(p_\text{colonize}\) than leaves lower
stomatal density but the same \gsmax. In other words, increasing \(D\)
and lowering \(S\) allows plants to reduce \(p_\text{colonize}\) while
maintaining \gsmax. The scaling relationship is slightly less than 1,
but still greater than 0.5, when \(H > 0\) (Figure \ref{fig:fig4}). In
this area of parameter space, lower stomatal density can reduce \fs{}
while \(p_\text{colonize}\) is constant, but this will still result in
lower \gsmax.

\begin{figure}
  \centering
    \includegraphics{../figures/fig4.pdf}
    \caption{Log-log scaling relationships between stomatal density ($D$, \textit{x}-axis) and size ($S$, \textit{y}-axis) that preserve the probability of colonization ($p_\text{colonize}$). In each panel, solid lines indicate values of $D$ and $S$ where $p_\text{colonize}$ is 0.025 (lowest line), 0.05, 0.1, 0.2, or 0.4 (highest line). For reference, dashed grey lines show scaling relationships that preserve \fs{} ($\beta = 1$, $\text{slope} = - 1 / \beta = -1$) and \gsmax{} ($\beta = 0.5$, $\text{slope} = - 1 / \beta = -2$) drawn through the centroid of the plotting region. When the death rate on the leaf surface is low ($H = 0$), the scaling exponent is very close to $\beta = 1$. When $H > 0$, $0.5 < \beta < 1$ and is slightly nonlinear on a log-log scale.}
    \label{fig:fig4}
\end{figure}

\hypertarget{discussion}{%
\section*{Discussion}\label{discussion}}
\addcontentsline{toc}{section}{Discussion}

Stomatal density and size set the upper limit on gas exchange in leaves
(Harrison et al., 2019) and is often closely related to operational
stomatal conductance in nature (Murray et al., 2019). Despite the fact
that many ecologically and economically significant plant pathogens
infect through stomata, the relationship between stomatal anatomy and
susceptibility to foliar pathogens is less clear than it is for gas
exchange. To develop testable predictions, we need mathematical models
that can clarify the potential for tradeoffs between stomatal
conductance, stomatal cover, and disease resistance. I used a spatially
explicit model of a pathogen searching for a stomate to colonize a host.
From this \protect\hyperlink{model}{Model}, I derived predictions about
the relationship between stomatal anatomy and disease resistance for the
first time. The model predicts that the probability of colonization is
not always proportional to the surface area of leaf covered by stomata
(\fs), as one might intuitively predict. If the leaf surface is a
hostile environment and pathogens have a limited time to search, lower
stomatal density decreases the probability of colonization even if \fs{}
is constant. However, \gsmax{} decreases proportionally more than the
probability of colonization. The model reveals the potential for
conflicting demands of maximizing disease resistance, minimizing
stomatal cover, and maintaining stomatal conductance. Including the
effect of anatomy on disease resistance therefore has the potential to
change our understading of how stomatal size-density scaling evolves in
land plants.

The model predicts that in most cases, increasing stomatal cover should
lead to a proportonal increase in susceptibility, which is the implicit
assumption of some empirical studies (e.g.~McKown et al. (2014); Tateda
et al. (2019); Dutton et al. (2019); Fetter et al. (2019)). It also
makes new, testable predictions that are less intuitive. At very low
\fs, there is a rapid increase in susceptibility (Figure
\ref{fig:fig3}a). If there are no stomata, the probability of
colonization is 0, so the first few stomata dramatically increase the
probability. This is unlikely to be significant for abaxial (lower) leaf
surfaces, which usually have most of the stomata (Salisbury, 1928;
Metcalfe and Chalk, 1950; Mott et al., 1984; Peat and Fitter, 1994;
Jordan et al., 2014; Muir, 2015; Bucher et al., 2017; Drake et al.,
2019). However, many adaxial (upper) leaf surfaces have zero or very few
stomata. Using adaxial leaf surfaces, it should be possible to test if
small changes in stomatal size or density have a larger effect on
disease susceptibility when \fs{} is low. The nonlinear increase in
\(p_\text{colonize}\) is less apparent when \(H > 0\) (Figure
\ref{fig:fig3}a). A more hostile microenvironment (e.g.~drier, higher
UV) should therefore reduce the effect of increased size or density as
low \fs. If true, the dimishing marginal effect of \fs{} on colonization
could explain why stomatal ratio on the upper and lower surface is
bimodal (Muir, 2015). The initial cost of adaxial (upper) stomata is
high, but if the benefits outweigh the costs, then equal stomatal
densities on each surface maximize CO\(_2\) supply for photosynthesis
(Parkhurst, 1978; Gutschick, 1984; Parkhurst and Mott, 1990).

An effect of stomatal size and density on susceptibility to foliar
pathogens could change our understanding of stomatal size-density
scaling. Since allocating leaf epidermis to stomata may be costly
(Franks and Farquhar, 2007; Assmann and Zeiger, 1987; Dow et al., 2014b;
Lehmann and Or, 2015; Baresch et al., 2019), selection should favor
leaves that achieve a desired \gsmax{} while minimizing \fs{} (de Boer
et al., 2016). Because of their different scaling exponents (Equation
\ref{eq:gsmax}, \ref{eq:fs}), smaller, densely packed stomata can
achieve the same \gsmax{} at minimum \fs. However, many leaves have
larger, sparsely packed stomata. Incorporating susceptibility to disease
may explain why. If pathogens have a limited time to find stomata before
dying (\(H > 0\)), then the scaling exponent between size and density
that keeps \(p_\text{colonize}\) constant is between 0.5 and 1, the
scaling exponents for \gsmax{} and \fs, respectively (Figure
\ref{fig:fig4}). Greater density of smaller stomata can increase
\gsmax{} while keeping \(p_\text{colonize}\) constant, but this will
increase \fs. Conversely, \fs{} could decrease while keeping
\(p_\text{colonize}\) constant, but this will decrease \gsmax. This sets
up the potential for conflict between competing goals. The optimal
stomatal size and density will therefore depend on the precise costs and
benefits of infection, stomatal conductance, and stomatal cover. This
may explain why many leaves have large, sparsely packed stomata despite
the fact that they could achieve the same \gsmax{} and lower \fs{} with
smaller, more densely packed stomata.

The model examines the probability of colonization for a single
pathogen. The calculated probabilities of colonization should not be
interpreted as exact predictions, but rather as depicting qualitative
relationships between stomatal anatomy and infection severity. The model
is most applicable to diseases where the host has some resistance. The
energetic cost and lost photosynthetic capacity (closed stomata,
necrosis, etc.) of dealing with a pathogen is assumed to be proportional
to the amount of infection. The actual fitness cost will be modulated by
the number of pathogens landing on the leaf and the cost of infection.
In environments with fewer or less virulent pathogens, the fitness cost
of infection will be less than in environments with more abundant,
virulent pathogens. The model is less relevant to very susceptible host
plants that can be severely damaged or killed by a small number of
colonizations that spread unchecked throughout the host tissue.

The purpose of this model is to provide a general foundation to examine
the relationship between stomatal size, density, and defense against
foliar pathogens. In its generality, it overlooks interesting natural
history and biologically important features of specific plants and their
pathogens. For example, some pathogens actively seek out and find
stomata (Kiefer et al., 2002), whereas I assumed that pathogens randomly
orient themselves on the leaf. Including sensing should increase the
probability of colonization. I also assumed that the plant does not
respond by, for example, closing stomata when it senses a pathogen.
However, plants can sense and close stomata when pathogens land on the
leaf, but pathogens can pry stomata back open (Melotto et al., 2006).
Future work on specific plant-pathogen interactions could build on this
model by adding more biological realism to provide more precise
predictions.

\hypertarget{conclusion}{%
\section*{Conclusion}\label{conclusion}}
\addcontentsline{toc}{section}{Conclusion}

The model makes two non-intuitive predictions. First, the effect of
increased stomatal density or size on susceptibility to foliar pathogens
is greatest when stomatal cover is very low. Second, maximizing disease
resistance sets up a potential conflict between minimizing stomatal
cover and maximizing stomatal conductance. The first prediction may be
relatively straightforward to test experimentally with adaxial (upper)
stomata that occur at low and moderate densities within the same or
closely related species (Muir et al., 2014; McKown et al., 2014; Fetter
et al., 2019). The second prediction about size-density scaling is more
complex because we would need to know the relationships between
colonization, stomatal cover, stomatal conductance, and fitness in
natural conditions. Testing these predictions in a variety of species
would help determine whether pathogens have played an important role
shaping stomatal anatomy in land plants.

\hypertarget{funding}{%
\section*{Funding}\label{funding}}
\addcontentsline{toc}{section}{Funding}

I am grateful startup funds from the University of Hawaii for supporting
this work.

\hypertarget{conflict-of-interest-statement}{%
\section*{Conflict of Interest
Statement}\label{conflict-of-interest-statement}}
\addcontentsline{toc}{section}{Conflict of Interest Statement}

The author declares that the research was conducted in the absence of
any commercial or financial relationships that could be construed as a
potential conflict of interest.

\hypertarget{supplementary-material}{%
\section*{Supplementary Material}\label{supplementary-material}}
\addcontentsline{toc}{section}{Supplementary Material}

I calculated \gsmax{} (Equation \ref{eq:gsmax}) to water vapor at a
reference leaf temperature (\(T_\mathrm{leaf} = 25 ^ \circ\) C)
following Sack and Buckley (2016). They defined a biophysical and
morphological constant as:

\begin{align*}
  b = & D_\text{wv} / v \\
  m = & \frac{\pi c ^ 2}{j ^ {0.5} (4 h j + \pi c)}
\end{align*}

\(b\) is the diffusion coefficient of water vapor in air
(\(D_\text{wv}\)) divided by the kinematic viscosity of dry air (\(v\)).
\(D_\text{wv} = 2.49 \times 10^{-5} ~ \text{m}^2 ~ \text{s}^{-1}\) and
\(v = 2.24 \times 10 ^ {-2} ~ \text{m}^3 ~ \text{mol} ^ {-1}\) at
\(25 ^ \circ\) (Monteith and Unsworth, 2013). For kidney-shaped guard
cells, \(c = h = j = 0.5\).

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-allaire_rmarkdown:_2019}{}%
Allaire, J., Xie, Y., McPherson, J., Luraschi, J., Ushey, K., Wickham,
H., Cheng, J., Chang, W., and Iannone, R. (2019). Rmarkdown: Dynamic
Documents for R. Available at:
\url{https://github.com/rstudio/rmarkdown}.

\leavevmode\hypertarget{ref-assmann_guard_1987}{}%
Assmann, S. M., and Zeiger, E. (1987). ``Guard Call Bioenergetics,'' in
\emph{Stomatal Function}, eds. E. Zeiger, G. D. Farquhar, and I. R.
Cowan (Stanford University Press), 163--193.

\leavevmode\hypertarget{ref-baresch_competition_2019}{}%
Baresch, A., Crifò, C., and Boyce, C. K. (2019). Competition for
epidermal space in the evolution of leaves with high physiological
rates. \emph{New Phytologist} 221, 628--639.
doi:\href{https://doi.org/10.1111/nph.15476}{10.1111/nph.15476}.

\leavevmode\hypertarget{ref-berry_stomata:_2010}{}%
Berry, J. A., Beerling, D. J., and Franks, P. J. (2010). Stomata: Key
players in the earth system, past and present. \emph{Current Opinion in
Plant Biology} 13, 232--239.
doi:\href{https://doi.org/10.1016/j.pbi.2010.04.013}{10.1016/j.pbi.2010.04.013}.

\leavevmode\hypertarget{ref-borchers_pracma:_2019}{}%
Borchers, H. W. (2019). Pracma: Practical Numerical Math Functions. R
package version 2.2.5. Available at:
\url{https://CRAN.R-project.org/package=pracma}.

\leavevmode\hypertarget{ref-brodribb_unified_2013}{}%
Brodribb, T. J., Jordan, G. J., and Carpenter, R. J. (2013). Unified
changes in cell size permit coordinated leaf evolution. \emph{New
Phytologist} 199, 559--570.
doi:\href{https://doi.org/10.1111/nph.12300}{10.1111/nph.12300}.

\leavevmode\hypertarget{ref-brown_static_1900}{}%
Brown, H. T., and Escombe, F. (1900). Static diffusion of gases and
liquids in relation to the assimilation of carbon and translocation in
plants. \emph{Proceedings of the Royal Society of London} 67, 124--128.

\leavevmode\hypertarget{ref-bucher_stomatal_2017}{}%
Bucher, S. F., Auerswald, K., Grün-Wenzel, C., Higgins, S. I., Garcia
Jorge, J., and Römermann, C. (2017). Stomatal traits relate to habitat
preferences of herbaceous species in a temperate climate. \emph{Flora}
229, 107--115.
doi:\href{https://doi.org/10.1016/j.flora.2017.02.011}{10.1016/j.flora.2017.02.011}.

\leavevmode\hypertarget{ref-buckley_how_2019}{}%
Buckley, T. N. (2019). How do stomata respond to water status? \emph{New
Phytologist} 224, 21--36.
doi:\href{https://doi.org/10.1111/nph.15899}{10.1111/nph.15899}.

\leavevmode\hypertarget{ref-chater_origins_2017}{}%
Chater, C. C. C., Caine, R. S., Fleming, A. J., and Gray, J. E. (2017).
Origins and Evolution of Stomatal Development. \emph{Plant Physiology}
174, 624--638.
doi:\href{https://doi.org/10.1104/pp.17.00183}{10.1104/pp.17.00183}.

\leavevmode\hypertarget{ref-de_boer_optimal_2016}{}%
de Boer, H. J., Price, C. A., Wagner-Cremer, F., Dekker, S. C., Franks,
P. J., and Veneklaas, E. J. (2016). Optimal allocation of leaf epidermal
area for gas exchange. \emph{New Phytologist} 210, 1219--1228.
doi:\href{https://doi.org/10.1111/nph.13929}{10.1111/nph.13929}.

\leavevmode\hypertarget{ref-dow_integrated_2014}{}%
Dow, G. J., Bergmann, D. C., and Berry, J. A. (2014a). An integrated
model of stomatal development and leaf physiology. \emph{New
Phytologist} 201, 1218--1226.

\leavevmode\hypertarget{ref-dow_physiological_2014}{}%
Dow, G. J., Berry, J. A., and Bergmann, D. C. (2014b). The physiological
importance of developmental mechanisms that enforce proper stomatal
spacing in \emph{Arabidopsis thaliana}. \emph{New Phytologist} 201,
1205--1217.
doi:\href{https://doi.org/10.1111/nph.12586}{10.1111/nph.12586}.

\leavevmode\hypertarget{ref-drake_two_2019}{}%
Drake, P. L., Boer, H. J., Schymanski, S. J., and Veneklaas, E. J.
(2019). Two sides to every leaf: Water and \textless span
style="font-variant:Small-caps;"\textgreater CO\textless/span\textgreater{}
\(_{\textrm{2}}\) transport in hypostomatous and amphistomatous leaves.
\emph{New Phytologist} 222, 1179--1187.
doi:\href{https://doi.org/10.1111/nph.15652}{10.1111/nph.15652}.

\leavevmode\hypertarget{ref-dutton_bacterial_2019}{}%
Dutton, C., Hõrak, H., Hepworth, C., Mitchell, A., Ton, J., Hunt, L.,
and Gray, J. E. (2019). Bacterial infection systemically suppresses
stomatal density. \emph{Plant, Cell \& Environment} 42, 2411--2421.
doi:\href{https://doi.org/10.1111/pce.13570}{10.1111/pce.13570}.

\leavevmode\hypertarget{ref-farquhar_stomatal_1982}{}%
Farquhar, G. D., and Sharkey, T. D. (1982). Stomatal Conductance and
Photosynthesis. \emph{Annual Review of Plant Physiology} 33, 317--345.
doi:\href{https://doi.org/10.1146/annurev.pp.33.060182.001533}{10.1146/annurev.pp.33.060182.001533}.

\leavevmode\hypertarget{ref-fawke_oomycete_2015}{}%
Fawke, S., Doumane, M., and Schornack, S. (2015). Oomycete Interactions
with Plants: Infection Strategies and Resistance Principles.
\emph{Microbiology and Molecular Biology Reviews} 79, 263--280.
doi:\href{https://doi.org/10.1128/MMBR.00010-15}{10.1128/MMBR.00010-15}.

\leavevmode\hypertarget{ref-fetter_trade-offs_2019}{}%
Fetter, K. C., Nelson, D. M., and Keller, S. R. (2019). Trade-offs and
selection conflicts in hybrid poplars indicate the stomatal ratio as an
important trait regulating disease resistance.
doi:\href{https://doi.org/10.1101/814046}{10.1101/814046}.

\leavevmode\hypertarget{ref-franks_co_2009}{}%
Franks, P. J., and Beerling, D. J. (2009a). CO \(_{\textrm{2}}\) -forced
evolution of plant gas exchange capacity and water-use efficiency over
the Phanerozoic. \emph{Geobiology} 7, 227--236.
doi:\href{https://doi.org/10.1111/j.1472-4669.2009.00193.x}{10.1111/j.1472-4669.2009.00193.x}.

\leavevmode\hypertarget{ref-franks_maximum_2009}{}%
Franks, P. J., and Beerling, D. J. (2009b). Maximum leaf conductance
driven by CO\(_{\textrm{2}}\) effects on stomatal size and density over
geologic time. \emph{Proceedings of the National Academy of Sciences}
106, 10343--10347.

\leavevmode\hypertarget{ref-franks_plasticity_2009}{}%
Franks, P. J., Drake, P. L., and Beerling, D. J. (2009). Plasticity in
maximum stomatal conductance constrained by negative correlation between
stomatal size and density: An analysis using \emph{Eucalyptus globulus}.
\emph{Plant, Cell \& Environment} 32, 1737--1748.
doi:\href{https://doi.org/10.1111/j.1365-3040.2009.002031.x}{10.1111/j.1365-3040.2009.002031.x}.

\leavevmode\hypertarget{ref-franks_effect_2001}{}%
Franks, P. J., and Farquhar, G. D. (2001). The effect of exogenous
abscisic acid on stomatal development, stomatal mechanics, and leaf gas
exchange in \emph{Tradescantia virginiana}. \emph{Plant Physiology} 125,
935--942.

\leavevmode\hypertarget{ref-franks_mechanical_2007}{}%
Franks, P. J., and Farquhar, G. D. (2007). The Mechanical Diversity of
Stomata and Its Significance in Gas-Exchange Control. \emph{Plant
Physiology} 143, 78--87.
doi:\href{https://doi.org/10.1104/pp.106.089367}{10.1104/pp.106.089367}.

\leavevmode\hypertarget{ref-franks_new_2014}{}%
Franks, P. J., Royer, D. L., Beerling, D. J., Van de Water, P. K.,
Cantrill, D. J., Barbour, M. M., and Berry, J. A. (2014). New
constraints on atmospheric CO \(_{\textrm{2}}\) concentration for the
Phanerozoic: Franks et al.: New constraints on Phanerozoic CO2.
\emph{Geophysical Research Letters} 41, 4685--4694.
doi:\href{https://doi.org/10.1002/2014GL060457}{10.1002/2014GL060457}.

\leavevmode\hypertarget{ref-gutschick_photosynthesis_1984}{}%
Gutschick, V. P. (1984). Photosynthesis model for C\(_{\textrm{3}}\)
leaves incorporating CO\(_{\textrm{2}}\) transport, propagation of
radiation, and biochemistry 1. Kinetics and their parameterization.
\emph{Photosynthetica} 18, 549--568.

\leavevmode\hypertarget{ref-harrison_influence_2019}{}%
Harrison, E. L., Arce Cubas, L., Gray, J. E., and Hepworth, C. (2019).
The influence of stomatal morphology and distribution on photosynthetic
gas exchange. \emph{The Plant Journal}, tpj.14560.
doi:\href{https://doi.org/10.1111/tpj.14560}{10.1111/tpj.14560}.

\leavevmode\hypertarget{ref-hetherington_role_2003}{}%
Hetherington, A. M., and Woodward, F. I. (2003). The role of stomata in
sensing and driving environmental change. \emph{Nature} 424, 901--908.
doi:\href{https://doi.org/10.1038/nature01843}{10.1038/nature01843}.

\leavevmode\hypertarget{ref-hoch_signaling_1987}{}%
Hoch, H. C., Staples, R. C., Whitehead, B., Comeau, J., and Wolf, E. D.
(1987). Signaling for Growth Orientation and Cell Differentiation by
Surface Topography in Uromyces. \emph{Science, New Series} 235,
1659--1662. Available at: \url{http://www.jstor.org/stable/1698314}.

\leavevmode\hypertarget{ref-jones_partitioning_1985}{}%
Jones, H. G. (1985). Partitioning stomatal and non-stomatal limitations
to photosynthesis. \emph{Plant, Cell \& Environment} 8, 95--104.
doi:\href{https://doi.org/10.1111/j.1365-3040.1985.tb01227.x}{10.1111/j.1365-3040.1985.tb01227.x}.

\leavevmode\hypertarget{ref-jordan_using_2014}{}%
Jordan, G. J., Carpenter, R. J., and Brodribb, T. J. (2014). Using
fossil leaves as evidence for open vegetation. \emph{Palaeogeography,
Palaeoclimatology, Palaeoecology} 395, 168--175.
doi:\href{https://doi.org/10.1016/j.palaeo.2013.12.035}{10.1016/j.palaeo.2013.12.035}.

\leavevmode\hypertarget{ref-kiefer_host_2002}{}%
Kiefer, B., Riemann, M., Büche, C., Kassemeyer, H.-H., and Nick, P.
(2002). The host guides morphogenesis and stomatal targeting in the
grapevine pathogen Plasmopara viticola. \emph{Planta} 215, 387--393.
doi:\href{https://doi.org/10.1007/s00425-002-0760-2}{10.1007/s00425-002-0760-2}.

\leavevmode\hypertarget{ref-lehmann_effects_2015}{}%
Lehmann, P., and Or, D. (2015). Effects of stomata clustering on leaf
gas exchange. \emph{New Phytologist} 207, 1015--1025.
doi:\href{https://doi.org/10.1111/nph.13442}{10.1111/nph.13442}.

\leavevmode\hypertarget{ref-mcelwain_using_2016}{}%
McElwain, J. C., Yiotis, C., and Lawson, T. (2016). Using modern plant
trait relationships between observed and theoretical maximum stomatal
conductance and vein density to examine patterns of plant
macroevolution. \emph{New Phytologist} 209, 94--103.
doi:\href{https://doi.org/10.1111/nph.13579}{10.1111/nph.13579}.

\leavevmode\hypertarget{ref-mckown_association_2014}{}%
McKown, A. D., Guy, R. D., Quamme, L., Klápště, J., La Mantia, J.,
Constabel, C. P., El-Kassaby, Y. A., Hamelin, R. C., Zifkin, M., and
Azam, M. S. (2014). Association genetics, geography and ecophysiology
link stomatal patterning in \emph{Populus trichocarpa} with carbon gain
and disease resistance trade-offs. \emph{Molecular Ecology} 23,
5771--5790.
doi:\href{https://doi.org/10.1111/mec.12969}{10.1111/mec.12969}.

\leavevmode\hypertarget{ref-mclachlan_gate_2014}{}%
McLachlan, D. H., Kopischke, M., and Robatzek, S. (2014). Gate control:
Guard cell regulation by microbial stress. \emph{New Phytologist} 203,
1049--1063.
doi:\href{https://doi.org/10.1111/nph.12916}{10.1111/nph.12916}.

\leavevmode\hypertarget{ref-melotto_plant_2006}{}%
Melotto, M., Underwood, W., Koczan, J., Nomura, K., and He, S. Y.
(2006). Plant Stomata Function in Innate Immunity against Bacterial
Invasion. \emph{Cell} 126, 969--980.
doi:\href{https://doi.org/10.1016/j.cell.2006.06.054}{10.1016/j.cell.2006.06.054}.

\leavevmode\hypertarget{ref-melotto_stomatal_2017}{}%
Melotto, M., Zhang, L., Oblessuc, P. R., and He, S. Y. (2017). Stomatal
Defense a Decade Later. \emph{Plant Physiology} 174, 561--571.
doi:\href{https://doi.org/10.1104/pp.16.01853}{10.1104/pp.16.01853}.

\leavevmode\hypertarget{ref-metcalfe_anatomy_1950}{}%
Metcalfe, C. R., and Chalk, L. (1950). \emph{Anatomy of the
dicotyledons, Vols. 1 \& 2}. First. Oxford: Oxford University Press.

\leavevmode\hypertarget{ref-meurer_sympy:_2017}{}%
Meurer, A., Smith, C. P., Paprocki, M., Čertík, O., Kirpichev, S. B.,
Rocklin, M., Kumar, A., Ivanov, S., Moore, J. K., Singh, S., et al.
(2017). SymPy: Symbolic computing in Python. \emph{PeerJ Computer
Science} 3, e103.
doi:\href{https://doi.org/10.7717/peerj-cs.103}{10.7717/peerj-cs.103}.

\leavevmode\hypertarget{ref-monteith_principles_2013}{}%
Monteith, J. L., and Unsworth, M. H. (2013). \emph{Principles of
environmental physics: Plants, animals, and the atmosphere}. 4th ed.
Amsterdam ; Boston: Elsevier/Academic Press.

\leavevmode\hypertarget{ref-morison_lateral_2005}{}%
Morison, J. I. L., Emily Gallouët, Lawson, T., Cornic, G., Herbin, R.,
and work(s): N. R. B. R. (2005). Lateral Diffusion of CO₂ in Leaves Is
Not Sufficient to Support Photosynthesis. \emph{Plant Physiology} 139,
254--266. Available at: \url{http://www.jstor.org/stable/4281859}.

\leavevmode\hypertarget{ref-mott_adaptive_1984}{}%
Mott, K. A., Gibson, A. C., and O'Leary, J. W. (1984). The adaptive
significance of amphistomatic leaves. \emph{Plant, Cell \& Environment}
5, 455--460.

\leavevmode\hypertarget{ref-muir_making_2015}{}%
Muir, C. D. (2015). Making pore choices: Repeated regime shifts in
stomatal ratio. \emph{Proceedings of the Royal Society B: Biological
Sciences} 282, 20151498.
doi:\href{https://doi.org/10.1098/rspb.2015.1498}{10.1098/rspb.2015.1498}.

\leavevmode\hypertarget{ref-muir_morphological_2014}{}%
Muir, C. D., Hangarter, R. P., Moyle, L. C., and Davis, P. A. (2014).
Morphological and anatomical determinants of mesophyll conductance in
wild relatives of tomato ( \emph{solanum} sect. \emph{Lycopersicon} ,
sect. \emph{Lycopersicoides} ; Solanaceae). \emph{Plant, Cell \&
Environment} 37, 1415--1426.
doi:\href{https://doi.org/10.1111/pce.12245}{10.1111/pce.12245}.

\leavevmode\hypertarget{ref-murray_consistent_2019}{}%
Murray, M., Soh, W. K., Yiotis, C., Spicer, R. A., Lawson, T., and
McElwain, J. C. (2019). Consistent relationship between field-measured
stomatal conductance and theoretical maximum stomatal conductance in
C\(_{\textrm{3}}\) woody angiosperms in four major biomes.
\emph{International Journal of Plant Sciences}, 706260.
doi:\href{https://doi.org/10.1086/706260}{10.1086/706260}.

\leavevmode\hypertarget{ref-murray_plant_2016}{}%
Murray, R. R., Emblow, M. S. M., Hetherington, A. M., and Foster, G. D.
(2016). Plant virus infections control stomatal development.
\emph{Scientific Reports} 6, 34507.
doi:\href{https://doi.org/10.1038/srep34507}{10.1038/srep34507}.

\leavevmode\hypertarget{ref-parkhurst_diffusion_1994}{}%
Parkhurst, D. F. (1994). Diffusion of
CO\textless sub\textgreater2\textless\textbackslash sub\textgreater{}
and Other Gases Inside Leaves. \emph{New Phytologist} 126, 449--479.
Available at: \url{http://www.jstor.org/stable/2557929}.

\leavevmode\hypertarget{ref-parkhurst_adaptive_1978}{}%
Parkhurst, D. F. (1978). The Adaptive Significance of Stomatal
Occurrence on One or Both Surfaces of Leaves. \emph{The Journal of
Ecology} 66, 367.
doi:\href{https://doi.org/10.2307/2259142}{10.2307/2259142}.

\leavevmode\hypertarget{ref-parkhurst_intercellular_1990}{}%
Parkhurst, D. F., and Mott, K. A. (1990). Intercellular Diffusion Limits
to CO \(_{\textrm{2}}\) Uptake in Leaves: Studies in Air and Helox.
\emph{Plant Physiology} 94, 1024--1032.
doi:\href{https://doi.org/10.1104/pp.94.3.1024}{10.1104/pp.94.3.1024}.

\leavevmode\hypertarget{ref-parlange_stomatal_1970}{}%
Parlange, J.-Y., and Waggoner, P. E. (1970). Stomatal Dimensions and
Resistance to Diffusion. \emph{Plant Physiology} 46, 337--342.
doi:\href{https://doi.org/10.1104/pp.46.2.337}{10.1104/pp.46.2.337}.

\leavevmode\hypertarget{ref-peat_comparative_1994}{}%
Peat, H. J., and Fitter, A. H. (1994). A comparative study of the
distribution and density of stomata in the British flora.
\emph{Biological Journal of the Linnean Society} 52, 377--393.

\leavevmode\hypertarget{ref-r_core_team_r:_2019}{}%
R Core Team (2019). \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing
Available at: \url{http://www.R-project.org/}.

\leavevmode\hypertarget{ref-sack_developmental_2016}{}%
Sack, L., and Buckley, T. N. (2016). The developmental basis of stomatal
density and flux. \emph{Plant Physiology}, pp.00476.2016.
doi:\href{https://doi.org/10.1104/pp.16.00476}{10.1104/pp.16.00476}.

\leavevmode\hypertarget{ref-sack_hydrology_2003}{}%
Sack, L., Cowan, P. D., Jaikumar, N., and Holbrook, N. M. (2003). The
'hydrology' of leaves: Co-ordination of structure and function in
temperate woody species. \emph{Plant, Cell and Environment} 26,
1343--1356.
doi:\href{https://doi.org/10.1046/j.0016-8025.2003.01058.x}{10.1046/j.0016-8025.2003.01058.x}.

\leavevmode\hypertarget{ref-salisbury_causes_1928}{}%
Salisbury, E. J. (1928). On the Causes and Ecological Significance of
Stomatal Frequency, with Special Reference to the Woodland Flora.
\emph{Philosophical Transactions of the Royal Society B: Biological
Sciences} 216, 1--65.
doi:\href{https://doi.org/10.1098/rstb.1928.0001}{10.1098/rstb.1928.0001}.

\leavevmode\hypertarget{ref-tateda_host_2019}{}%
Tateda, C., Obara, K., Abe, Y., Sekine, R., Nekoduka, S., Hikage, T.,
Nishihara, M., Sekine, K.-T., and Fujisaki, K. (2019). The Host Stomatal
Density Determines Resistance to \emph{Septoria gentianae} in Japanese
Gentian. \emph{Molecular Plant-Microbe Interactions} 32, 428--436.
doi:\href{https://doi.org/10.1094/MPMI-05-18-0114-R}{10.1094/MPMI-05-18-0114-R}.

\leavevmode\hypertarget{ref-ticha_photosynthetic_1982}{}%
Tichá, I. (1982). Photosynthetic characteristics during ontogenesis of
leaves 7. Stomata density and sizes. \emph{Photosynthetica} 16,
375--471.

\leavevmode\hypertarget{ref-underwood_role_2007}{}%
Underwood, W., Melotto, M., and He, S. Y. (2007). Role of plant stomata
in bacterial invasion. \emph{Cellular Microbiology} 9, 1621--1629.
doi:\href{https://doi.org/10.1111/j.1462-5822.2007.00938.x}{10.1111/j.1462-5822.2007.00938.x}.

\leavevmode\hypertarget{ref-weiss_untersuchungen_1865}{}%
Weiss, A. (1865). Untersuchungen über die Zahlen- und
Grössenverhältnisse der Spaltöffnungen. \emph{Jahrbücher für
Wissenschaftliche Botanik} 4, 125--196.

\leavevmode\hypertarget{ref-xie_r_2018}{}%
Xie, Y., Allaire, J. J., and Grolemund, G. (2018). \emph{R Markdown: The
definitive guide}. Boca Raton: Taylor \& Francis, CRC Press.

\leavevmode\hypertarget{ref-zeng_plant_2010}{}%
Zeng, W., Melotto, M., and He, S. Y. (2010). Plant stomata: A checkpoint
of host immunity and pathogen virulence. \emph{Current Opinion in
Biotechnology} 21, 599--603.
doi:\href{https://doi.org/10.1016/j.copbio.2010.05.006}{10.1016/j.copbio.2010.05.006}.

\end{document}