### 2020-07-13 ms/appendix2.tex
r1

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is just an example/guide for you to refer to when submitting manuscripts to Frontiers, it is not mandatory to use Frontiers .cls files nor frontiers.tex  %
% This will only generate the Manuscript, the final article will be typeset by Frontiers after acceptance.
%                                              %
%                                                                                                                                                         %
% When submitting your files, remember to upload this *tex file, the pdf generated with it, the *bib file (if bibliography is not within the *tex) and all the figures.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Version 3.4 Generated 2018/06/15 %%%
%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%

\documentclass[utf8]{frontiersSCNS}

%\setcitestyle{square} % for Physics and Applied Mathematics and Statistics articles
\usepackage{url,hyperref,lineno,microtype,subcaption}
\usepackage[onehalfspacing]{setspace}

\linenumbers

% BELOW TAKEN FROM rticles plos template
%
% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% hyperref package, useful for hyperlinks
\usepackage{hyperref}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% Sweave(-like)
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
\newenvironment{Schunk}{}{}
\DefineVerbatimEnvironment{Code}{Verbatim}{}
\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
\newenvironment{CodeChunk}{}{}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Below is from frontiers
%
\bibliographystyle{frontiersinSCNS}
% Use doublespacing - comment out for single spacing
%\usepackage{setspace}
%\doublespacing

% Leave a blank line between paragraphs instead of using \\

\def\keyFont{\fontsize{8}{11}\helveticabold }

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW

%% END MACROS SECTION

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{soul}
\usepackage{tikz}

% short-cut for fs, gmax, and gop
\newcommand{\fs}{$f_\text{S}$}
\newcommand{\gsmax}{$g_\text{s,max}$}
\newcommand{\gsop}{$g_\text{s,op}$}

\def\Authors{
  Christopher D. Muir\,\textsuperscript{1*}}

\def\Address{

  \textsuperscript{1} School of Life Sciences, University of Hawaii,  Honolulu,  Hawaii,  USA
  }

  \def\corrAuthor{Christopher D. Muir}\def\corrAddress{,  }\def\corrEmail{\href{mailto:cdmuir@hawaii.edu}{\nolinkurl{cdmuir@hawaii.edu}}}
  \def\firstAuthorLast{Muir {et~al.}}

\begin{document}
\onecolumn
\firstpage{1}

\title[]{Appendix 2: Additional R packages}
\author[\firstAuthorLast]{\Authors}
\address{} %This field will be automatically populated
\correspondance{} %This field will be automatically populated

\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}

\maketitle

\begin{longtable}[]{@{}lll@{}}
\toprule
package & version & reference\tabularnewline
\midrule
\endhead
assertthat & 0.2.1 & Wickham (2019a)\tabularnewline
bibtex & 0.4.2.2 & Francois (2020)\tabularnewline
BiocManager & 1.30.10 & Morgan (2019)\tabularnewline
cli & 2.0.2 & Cs\a'ardi (2020)\tabularnewline
codetools & 0.2-16 & Tierney (2018)\tabularnewline
colorspace & 1.4-1 & Zeileis et al. (2019)\tabularnewline
& & Zeileis et al. (2009)\tabularnewline
& & Stauffer et al. (2009)\tabularnewline
cowplot & 1.0.0 & Wilke (2019)\tabularnewline
crayon & 1.3.4 & Cs\a'ardi (2017)\tabularnewline
digest & 0.6.25 & Antoine Lucas et al. (2020)\tabularnewline
dplyr & 1.0.0 & Wickham et al. (2020a)\tabularnewline
ellipsis & 0.3.1 & Wickham (2020)\tabularnewline
evaluate & 0.14 & Wickham and Xie (2019)\tabularnewline
fansi & 0.4.1 & Gaslam (2020)\tabularnewline
farver & 2.0.3 & Pedersen et al. (2020)\tabularnewline
filehash & 2.4-2 & Peng (2006)\tabularnewline
furrr & 0.1.0 & Vaughan and Dancho (2018)\tabularnewline
future & 1.18.0 & Bengtsson (2020)\tabularnewline
generics & 0.0.2 & Kuhn et al. (2018)\tabularnewline
ggforce & 0.3.2 & Pedersen (2020)\tabularnewline
ggimage & 0.2.8 & Yu (2020a)\tabularnewline
gginnards & 0.0.3 & Aphalo (2019)\tabularnewline
ggplot2 & 3.3.2 & Wickham (2016)\tabularnewline
ggplotify & 0.0.5 & Yu (2020b)\tabularnewline
globals & 0.12.5 & Bengtsson (2019a)\tabularnewline
glue & 1.4.1 & Hester (2020)\tabularnewline
gridGraphics & 0.5-0 & Murrell and Wen (2020)\tabularnewline
gtable & 0.3.0 & Wickham and Pedersen (2019)\tabularnewline
hms & 0.5.3 & Müller (2020)\tabularnewline
htmltools & 0.5.0 & Cheng et al. (2020)\tabularnewline
httr & 1.4.1 & Wickham (2019b)\tabularnewline
jsonlite & 1.7.0 & Ooms (2014)\tabularnewline
knitcitations & 1.0.10 & Boettiger (2019)\tabularnewline
knitr & 1.29 & Xie (2020a)\tabularnewline
& & Xie (2015)\tabularnewline
& & Xie (2014)\tabularnewline
lifecycle & 0.2.0 & Henry (2020)\tabularnewline
listenv & 0.8.0 & Bengtsson (2019b)\tabularnewline
lubridate & 1.7.9 & Grolemund and Wickham (2011)\tabularnewline
magick & 2.4.0 & Ooms (2020)\tabularnewline
magrittr & 1.5 & Bache and Wickham (2014)\tabularnewline
MASS & 7.3-51.6 & Venables and Ripley (2002)\tabularnewline
munsell & 0.5.0 & Wickham (2018)\tabularnewline
pillar & 1.4.6 & Müller and Wickham (2020a)\tabularnewline
pkgconfig & 2.0.3 & Cs\a'ardi (2019)\tabularnewline
plyr & 1.8.6 & Wickham (2011)\tabularnewline
polyclip & 1.10-0 & Johnson and Baddeley (2019)\tabularnewline
pracma & 2.2.9 & Borchers (2019)\tabularnewline
purrr & 0.3.4 & Henry and Wickham (2020a)\tabularnewline
R6 & 2.4.1 & Chang (2019)\tabularnewline
Rcpp & 1.0.5 & Eddelbuettel and François (2011)\tabularnewline
& & Eddelbuettel (2013)\tabularnewline
& & Eddelbuettel and Balamuta (2017)\tabularnewline
readr & 1.3.1 & Wickham et al. (2018)\tabularnewline
RefManageR & 1.2.12 & McLean (2017)\tabularnewline
& & McLean (2014)\tabularnewline
rlang & 0.4.7 & Henry and Wickham (2020b)\tabularnewline
rmarkdown & 2.3 & Allaire et al. (2020a)\tabularnewline
& & Xie et al. (2018)\tabularnewline
rticles & 0.14 & Allaire et al. (2020b)\tabularnewline
rvcheck & 0.1.8 & Yu (2020c)\tabularnewline
scales & 1.1.1 & Wickham and Seidel (2020)\tabularnewline
sessioninfo & 1.1.1 & Cs\a'ardi et al. (2018)\tabularnewline
stringi & 1.4.6 & Gagolewski (2020)\tabularnewline
stringr & 1.4.0 & Wickham (2019c)\tabularnewline
tibble & 3.0.3 & Müller and Wickham (2020b)\tabularnewline
tidyr & 1.1.0 & Wickham and Henry (2020)\tabularnewline
tidyselect & 1.1.0 & Henry and Wickham (2020c)\tabularnewline
tikzDevice & 0.12.3.1 & Sharpsteen and Bracken (2020)\tabularnewline
tweenr & 1.0.1 & Pedersen (2018)\tabularnewline
units & 0.6-7 & Pebesma et al. (2016)\tabularnewline
vctrs & 0.3.1 & Wickham et al. (2020b)\tabularnewline
withr & 2.2.0 & Hester et al. (2020)\tabularnewline
xfun & 0.15 & Xie (2020b)\tabularnewline
xml2 & 1.3.2 & Wickham et al. (2020c)\tabularnewline
yaml & 2.2.1 & Stephens et al. (2020)\tabularnewline
\bottomrule
\end{longtable}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-allaire2020rmarkdown}{}%
Allaire, J., Xie, Y., McPherson, J., Luraschi, J., Ushey, K., Atkins,
A., Wickham, H., Cheng, J., Chang, W., and Iannone, R. (2020a).
\emph{Rmarkdown: Dynamic documents for r}. Available at:
\url{https://github.com/rstudio/rmarkdown}.

\leavevmode\hypertarget{ref-allaire2020rticles}{}%
Allaire, J., Xie, Y., R Foundation, Wickham, H., Journal of Statistical
Software, Vaidyanathan, R., Association for Computing Machinery,
Boettiger, C., Elsevier, Broman, K., et al. (2020b). \emph{Rticles:
Article formats for r markdown}. Available at:
\url{https://CRAN.R-project.org/package=rticles}.

\leavevmode\hypertarget{ref-lucas2020digest}{}%
Antoine Lucas, D. E. with contributions by, Tuszynski, J., Bengtsson,
H., Urbanek, S., Frasca, M., Lewis, B., Stokely, M., Muehleisen, H.,
Murdoch, D., Hester, J., et al. (2020). \emph{Digest: Create compact
hash digests of r objects}. Available at:
\url{https://CRAN.R-project.org/package=digest}.

\leavevmode\hypertarget{ref-aphalo2019gginnards}{}%
Aphalo, P. J. (2019). \emph{Gginnards: Explore the innards of 'ggplot2'
objects}. Available at:
\url{https://CRAN.R-project.org/package=gginnards}.

\leavevmode\hypertarget{ref-bache2014magrittr}{}%
Bache, S. M., and Wickham, H. (2014). \emph{Magrittr: A forward-pipe
operator for r}. Available at:
\url{https://CRAN.R-project.org/package=magrittr}.

\leavevmode\hypertarget{ref-bengtsson2020future}{}%
Bengtsson, H. (2020). \emph{Future: Unified parallel and distributed
processing in r for everyone}. Available at:
\url{https://CRAN.R-project.org/package=future}.

\leavevmode\hypertarget{ref-bengtsson2019globals}{}%
Bengtsson, H. (2019a). \emph{Globals: Identify global objects in r
expressions}. Available at:
\url{https://CRAN.R-project.org/package=globals}.

\leavevmode\hypertarget{ref-bengtsson2019listenv}{}%
Bengtsson, H. (2019b). \emph{Listenv: Environments behaving (almost) as
lists}. Available at: \url{https://CRAN.R-project.org/package=listenv}.

\leavevmode\hypertarget{ref-boettiger2019knitcitations}{}%
Boettiger, C. (2019). \emph{Knitcitations: Citations for 'knitr'
markdown files}. Available at:
\url{https://CRAN.R-project.org/package=knitcitations}.

\leavevmode\hypertarget{ref-borchers2019pracma}{}%
Borchers, H. W. (2019). \emph{Pracma: Practical numerical math
functions}. Available at:
\url{https://CRAN.R-project.org/package=pracma}.

\leavevmode\hypertarget{ref-chang2019encapsulated}{}%
Chang, W. (2019). \emph{R6: Encapsulated classes with reference
semantics}. Available at: \url{https://CRAN.R-project.org/package=R6}.

\leavevmode\hypertarget{ref-cheng2020htmltools}{}%
Cheng, J., Sievert, C., Chang, W., Xie, Y., and Allen, J. (2020).
\emph{Htmltools: Tools for html}. Available at:
\url{https://CRAN.R-project.org/package=htmltools}.

\leavevmode\hypertarget{ref-csrdi2020helpers}{}%
Cs\a'ardi, G. (2020). \emph{Cli: Helpers for developing command line
interfaces}. Available at: \url{https://CRAN.R-project.org/package=cli}.

\leavevmode\hypertarget{ref-csrdi2017crayon}{}%
Cs\a'ardi, G. (2017). \emph{Crayon: Colored terminal output}. Available
at: \url{https://CRAN.R-project.org/package=crayon}.

\leavevmode\hypertarget{ref-csrdi2019pkgconfig}{}%
Cs\a'ardi, G. (2019). \emph{Pkgconfig: Private configuration for 'r'
packages}. Available at:
\url{https://CRAN.R-project.org/package=pkgconfig}.

\leavevmode\hypertarget{ref-csrdi2018sessioninfo}{}%
Cs\a'ardi, G., core, R., Wickham, H., Chang, W., Flight, R. M., Müller,
K., and Hester, J. (2018). \emph{Sessioninfo: R session information}.
Available at: \url{https://CRAN.R-project.org/package=sessioninfo}.

\leavevmode\hypertarget{ref-eddelbuettel2013seamless}{}%
Eddelbuettel, D. (2013). \emph{Seamless R and C++ integration with
Rcpp}. New York: Springer
doi:\href{https://doi.org/10.1007/978-1-4614-6868-4}{10.1007/978-1-4614-6868-4}.

\leavevmode\hypertarget{ref-eddelbuettel2017extending}{}%
Eddelbuettel, D., and Balamuta, J. J. (2017). Extending extitR with
extitC++: A Brief Introduction to extitRcpp. \emph{PeerJ Preprints} 5,
e3188v1.
doi:\href{https://doi.org/10.7287/peerj.preprints.3188v1}{10.7287/peerj.preprints.3188v1}.

\leavevmode\hypertarget{ref-eddelbuettel2011rcpp}{}%
Eddelbuettel, D., and François, R. (2011). Rcpp: Seamless R and C++
integration. \emph{Journal of Statistical Software} 40, 1--18.
doi:\href{https://doi.org/10.18637/jss.v040.i08}{10.18637/jss.v040.i08}.

\leavevmode\hypertarget{ref-francois2020bibtex}{}%
Francois, R. (2020). \emph{Bibtex: Bibtex parser}. Available at:
\url{https://CRAN.R-project.org/package=bibtex}.

\leavevmode\hypertarget{ref-gagolewski2020package}{}%
Gagolewski, M. (2020). \emph{R package stringi: Character string
processing facilities}. Available at:
\url{http://www.gagolewski.com/software/stringi/}.

\leavevmode\hypertarget{ref-gaslam2020fansi}{}%
Gaslam, B. (2020). \emph{Fansi: ANSI control sequence aware string
functions}. Available at:
\url{https://CRAN.R-project.org/package=fansi}.

\leavevmode\hypertarget{ref-grolemund2011dates}{}%
Grolemund, G., and Wickham, H. (2011). Dates and times made easy with
lubridate. \emph{Journal of Statistical Software} 40, 1--25. Available
at: \url{http://www.jstatsoft.org/v40/i03/}.

\leavevmode\hypertarget{ref-henry2020lifecycle}{}%
Henry, L. (2020). \emph{Lifecycle: Manage the life cycle of your package
functions}. Available at:
\url{https://CRAN.R-project.org/package=lifecycle}.

\leavevmode\hypertarget{ref-henry2020purrr}{}%
Henry, L., and Wickham, H. (2020a). \emph{Purrr: Functional programming
tools}. Available at: \url{https://CRAN.R-project.org/package=purrr}.

\leavevmode\hypertarget{ref-henry2020rlang}{}%
Henry, L., and Wickham, H. (2020b). \emph{Rlang: Functions for base
types and core r and 'tidyverse' features}. Available at:
\url{https://CRAN.R-project.org/package=rlang}.

\leavevmode\hypertarget{ref-henry2020tidyselect}{}%
Henry, L., and Wickham, H. (2020c). \emph{Tidyselect: Select from a set
of strings}. Available at:
\url{https://CRAN.R-project.org/package=tidyselect}.

\leavevmode\hypertarget{ref-hester2020glue}{}%
Hester, J. (2020). \emph{Glue: Interpreted string literals}. Available
at: \url{https://CRAN.R-project.org/package=glue}.

\leavevmode\hypertarget{ref-hester2020withr}{}%
Hester, J., Müller, K., Ushey, K., Wickham, H., and Chang, W. (2020).
\emph{Withr: Run code 'with' temporarily modified global state}.
Available at: \url{https://CRAN.R-project.org/package=withr}.

\leavevmode\hypertarget{ref-johnson2019polyclip}{}%
Johnson, A., and Baddeley, A. (2019). \emph{Polyclip: Polygon clipping}.
Available at: \url{https://CRAN.R-project.org/package=polyclip}.

\leavevmode\hypertarget{ref-kuhn2018generics}{}%
Kuhn, M., Wickham, H., and Vaughan, D. (2018). \emph{Generics: Common s3
generics not provided by base r methods related to model fitting}.
Available at: \url{https://CRAN.R-project.org/package=generics}.

\leavevmode\hypertarget{ref-mclean2017refmanager}{}%
McLean, M. W. (2017). RefManageR: Import and manage bibtex and biblatex
references in r. \emph{The Journal of Open Source Software}.
doi:\href{https://doi.org/10.21105/joss.00338}{10.21105/joss.00338}.

\leavevmode\hypertarget{ref-mclean2014straightforward}{}%
McLean, M. W. (2014). \emph{Straightforward bibliography management in r
using the refmanager package}. Available at:
\url{https://arxiv.org/abs/1403.2036}.

\leavevmode\hypertarget{ref-morgan2019biocmanager}{}%
Morgan, M. (2019). \emph{BiocManager: Access the bioconductor project
package repository}. Available at:
\url{https://CRAN.R-project.org/package=BiocManager}.

\leavevmode\hypertarget{ref-murrell2020gridgraphics}{}%
Murrell, P., and Wen, Z. (2020). \emph{GridGraphics: Redraw base
graphics using 'grid' graphics}. Available at:
\url{https://CRAN.R-project.org/package=gridGraphics}.

\leavevmode\hypertarget{ref-mller2020pretty}{}%
Müller, K. (2020). \emph{Hms: Pretty time of day}. Available at:
\url{https://CRAN.R-project.org/package=hms}.

\leavevmode\hypertarget{ref-mller2020pillar}{}%
Müller, K., and Wickham, H. (2020a). \emph{Pillar: Coloured formatting
for columns}. Available at:
\url{https://CRAN.R-project.org/package=pillar}.

\leavevmode\hypertarget{ref-mller2020tibble}{}%
Müller, K., and Wickham, H. (2020b). \emph{Tibble: Simple data frames}.
Available at: \url{https://CRAN.R-project.org/package=tibble}.

\leavevmode\hypertarget{ref-ooms2020magick}{}%
Ooms, J. (2020). \emph{Magick: Advanced graphics and image-processing in
r}. Available at: \url{https://CRAN.R-project.org/package=magick}.

\leavevmode\hypertarget{ref-ooms2014jsonlite}{}%
Ooms, J. (2014). The jsonlite package: A practical and consistent
mapping between json data and r objects. \emph{arXiv:1403.2805
{[}stat.CO{]}}. Available at: \url{https://arxiv.org/abs/1403.2805}.

\leavevmode\hypertarget{ref-pebesma2016measurement}{}%
Pebesma, E., Mailund, T., and Hiebert, J. (2016). Measurement units in
R. \emph{R Journal} 8, 486--494.
doi:\href{https://doi.org/10.32614/RJ-2016-061}{10.32614/RJ-2016-061}.

\leavevmode\hypertarget{ref-pedersen2020ggforce}{}%
Pedersen, T. L. (2020). \emph{Ggforce: Accelerating 'ggplot2'}.
Available at: \url{https://CRAN.R-project.org/package=ggforce}.

\leavevmode\hypertarget{ref-pedersen2018tweenr}{}%
Pedersen, T. L. (2018). \emph{Tweenr: Interpolate data for smooth
animations}. Available at:
\url{https://CRAN.R-project.org/package=tweenr}.

\leavevmode\hypertarget{ref-pedersen2020farver}{}%
Pedersen, T. L., Nicolae, B., and François, R. (2020). \emph{Farver:
High performance colour space manipulation}. Available at:
\url{https://CRAN.R-project.org/package=farver}.

\leavevmode\hypertarget{ref-peng2006interacting}{}%
Peng, R. D. (2006). Interacting with data using the filehash package.
\emph{R News} 6, 19--24. Available at:
\url{https://cran.r-project.org/doc/Rnews/}.

\leavevmode\hypertarget{ref-sharpsteen2020tikzdevice}{}%
Sharpsteen, C., and Bracken, C. (2020). \emph{TikzDevice: R graphics
output in latex format}. Available at:
\url{https://CRAN.R-project.org/package=tikzDevice}.

\leavevmode\hypertarget{ref-stauffer2009somewhere}{}%
Stauffer, R., Mayr, G. J., Dabernig, M., and Zeileis, A. (2009).
Somewhere over the rainbow: How to make effective use of colors in
meteorological visualizations. \emph{Bulletin of the American
Meteorological Society} 96, 203--216.
doi:\href{https://doi.org/10.1175/BAMS-D-13-00155.1}{10.1175/BAMS-D-13-00155.1}.

\leavevmode\hypertarget{ref-stephens2020yaml}{}%
Stephens, J., Simonov, K., Xie, Y., Dong, Z., Wickham, H., Horner, J.,
reikoch, Beasley, W., O'Connor, B., and Warnes, G. R. (2020).
\emph{Yaml: Methods to convert r data to yaml and back}. Available at:
\url{https://CRAN.R-project.org/package=yaml}.

\leavevmode\hypertarget{ref-tierney2018codetools}{}%
Tierney, L. (2018). \emph{Codetools: Code analysis tools for r}.
Available at: \url{https://CRAN.R-project.org/package=codetools}.

\leavevmode\hypertarget{ref-vaughan2018furrr}{}%
Vaughan, D., and Dancho, M. (2018). \emph{Furrr: Apply mapping functions
in parallel using futures}. Available at:
\url{https://CRAN.R-project.org/package=furrr}.

\leavevmode\hypertarget{ref-venables2002modern}{}%
Venables, W. N., and Ripley, B. D. (2002). \emph{Modern applied
statistics with s}. Fourth. New York: Springer Available at:
\url{http://www.stats.ox.ac.uk/pub/MASS4}.

\leavevmode\hypertarget{ref-wickham2018munsell}{}%
Wickham, C. (2018). \emph{Munsell: Utilities for using munsell colours}.
Available at: \url{https://CRAN.R-project.org/package=munsell}.

\leavevmode\hypertarget{ref-wickham2019assertthat}{}%
Wickham, H. (2019a). \emph{Assertthat: Easy pre and post assertions}.
Available at: \url{https://CRAN.R-project.org/package=assertthat}.

\leavevmode\hypertarget{ref-wickham2020ellipsis}{}%
Wickham, H. (2020). \emph{Ellipsis: Tools for working with ...}
Available at: \url{https://CRAN.R-project.org/package=ellipsis}.

\leavevmode\hypertarget{ref-wickham2016ggplot2}{}%
Wickham, H. (2016). \emph{Ggplot2: Elegant graphics for data analysis}.
Springer-Verlag New York Available at:
\url{https://ggplot2.tidyverse.org}.

\leavevmode\hypertarget{ref-wickham2019httr}{}%
Wickham, H. (2019b). \emph{Httr: Tools for working with urls and http}.
Available at: \url{https://CRAN.R-project.org/package=httr}.

\leavevmode\hypertarget{ref-wickham2019stringr}{}%
Wickham, H. (2019c). \emph{Stringr: Simple, consistent wrappers for
common string operations}. Available at:
\url{https://CRAN.R-project.org/package=stringr}.

\leavevmode\hypertarget{ref-wickham2011split}{}%
Wickham, H. (2011). The split-apply-combine strategy for data analysis.
\emph{Journal of Statistical Software} 40, 1--29. Available at:
\url{http://www.jstatsoft.org/v40/i01/}.

\leavevmode\hypertarget{ref-wickham2020dplyr}{}%
Wickham, H., François, R., Henry, L., and Müller, K. (2020a).
\emph{Dplyr: A grammar of data manipulation}. Available at:
\url{https://CRAN.R-project.org/package=dplyr}.

\leavevmode\hypertarget{ref-wickham2020tidyr}{}%
Wickham, H., and Henry, L. (2020). \emph{Tidyr: Tidy messy data}.
Available at: \url{https://CRAN.R-project.org/package=tidyr}.

\leavevmode\hypertarget{ref-wickham2020vctrs}{}%
Wickham, H., Henry, L., and Vaughan, D. (2020b). \emph{Vctrs: Vector
helpers}. Available at: \url{https://CRAN.R-project.org/package=vctrs}.

\leavevmode\hypertarget{ref-wickham2018readr}{}%
Wickham, H., Hester, J., and Francois, R. (2018). \emph{Readr: Read
rectangular text data}. Available at:
\url{https://CRAN.R-project.org/package=readr}.

\leavevmode\hypertarget{ref-wickham2020xml2}{}%
Wickham, H., Hester, J., and Ooms, J. (2020c). \emph{Xml2: Parse xml}.
Available at: \url{https://CRAN.R-project.org/package=xml2}.

\leavevmode\hypertarget{ref-wickham2019gtable}{}%
Wickham, H., and Pedersen, T. L. (2019). \emph{Gtable: Arrange 'grobs'
in tables}. Available at:
\url{https://CRAN.R-project.org/package=gtable}.

\leavevmode\hypertarget{ref-wickham2020scales}{}%
Wickham, H., and Seidel, D. (2020). \emph{Scales: Scale functions for
visualization}. Available at:
\url{https://CRAN.R-project.org/package=scales}.

\leavevmode\hypertarget{ref-wickham2019evaluate}{}%
Wickham, H., and Xie, Y. (2019). \emph{Evaluate: Parsing and evaluation
tools that provide more details than the default}. Available at:
\url{https://CRAN.R-project.org/package=evaluate}.

\leavevmode\hypertarget{ref-wilke2019cowplot}{}%
Wilke, C. O. (2019). \emph{Cowplot: Streamlined plot theme and plot
annotations for 'ggplot2'}. Available at:
\url{https://CRAN.R-project.org/package=cowplot}.

\leavevmode\hypertarget{ref-xie2015dynamic}{}%
Xie, Y. (2015). \emph{Dynamic documents with R and knitr}. 2nd ed. Boca
Raton, Florida: Chapman; Hall/CRC Available at:
\url{https://yihui.org/knitr/}.

\leavevmode\hypertarget{ref-xie2014knitr}{}%
Xie, Y. (2014). ``Knitr: A comprehensive tool for reproducible research
in R,'' in \emph{Implementing reproducible computational research}, eds.
V. Stodden, F. Leisch, and R. D. Peng (Chapman; Hall/CRC). Available at:
\url{http://www.crcpress.com/product/isbn/9781466561595}.

\leavevmode\hypertarget{ref-xie2020knitr}{}%
Xie, Y. (2020a). \emph{Knitr: A general-purpose package for dynamic
report generation in r}. Available at: \url{https://yihui.org/knitr/}.

\leavevmode\hypertarget{ref-xie2020xfun}{}%
Xie, Y. (2020b). \emph{Xfun: Miscellaneous functions by 'yihui xie'}.
Available at: \url{https://CRAN.R-project.org/package=xfun}.

\leavevmode\hypertarget{ref-xie2018markdown}{}%
Xie, Y., Allaire, J. J., and Grolemund, G. (2018). \emph{R markdown: The
definitive guide}. Boca Raton, Florida: Chapman; Hall/CRC Available at:
\url{https://bookdown.org/yihui/rmarkdown}.

\leavevmode\hypertarget{ref-yu2020ggimage}{}%
Yu, G. (2020a). \emph{Ggimage: Use image in 'ggplot2'}. Available at:
\url{https://CRAN.R-project.org/package=ggimage}.

\leavevmode\hypertarget{ref-yu2020ggplotify}{}%
Yu, G. (2020b). \emph{Ggplotify: Convert plot to 'grob' or 'ggplot'
object}. Available at:
\url{https://CRAN.R-project.org/package=ggplotify}.

\leavevmode\hypertarget{ref-yu2020rvcheck}{}%
Yu, G. (2020c). \emph{Rvcheck: R/package version check}. Available at:
\url{https://CRAN.R-project.org/package=rvcheck}.

\leavevmode\hypertarget{ref-zeileis2019colorspace}{}%
Zeileis, A., Fisher, J. C., Hornik, K., Ihaka, R., McWhite, C. D.,
Murrell, P., Stauffer, R., and Wilke, C. O. (2019). colorspace: A
toolbox for manipulating and assessing colors and palettes. arXiv.org
E-Print Archive Available at: \url{http://arxiv.org/abs/1903.06490}.

\leavevmode\hypertarget{ref-zeileis2009escaping}{}%
Zeileis, A., Hornik, K., and Murrell, P. (2009). Escaping RGBland:
Selecting colors for statistical graphics. \emph{Computational
Statistics \& Data Analysis} 53, 3259--3270.
doi:\href{https://doi.org/10.1016/j.csda.2008.11.033}{10.1016/j.csda.2008.11.033}.

\end{document}